---
title: "BNLL hydrophysiology analyses"
author: "Savannah Weaver"
output: pdf_document
toc: TRUE
---


# Packages

```{r setup, echo=T, results='hide', message=FALSE}
if (!require("tidyverse")) install.packages("tidyverse") 
library("tidyverse")# for dplyr workflow of calculations
if (!require("UsingR")) install.packages("UsingR") 
library("UsingR")
if (!require("MASS")) install.packages("MASS") 
library("MASS")
if (!require("lme4")) install.packages("lme4") 
library("lme4")
if (!require("lmerTest")) install.packages("lmerTest") 
library("lmerTest")
```


# Load in Data

This data was collected during fieldwork with Blunt-nosed Leopard Lizards (*Gambelia sila*) in the Elkhorn Plain of the Carrizo Plain National Monument in California, USA during the 2021 active season (April-August). See published paper (doi) for complete methods and discussion of these analyses. 


## Collared Lizards

The first dataset loaded in is for lizards we radio-collared and tracked throughout the summer.

```{r get data collared lizards}
lizards_collared <- read.csv("data/collared_measurements.csv") %>%
                
                # format data classes
  dplyr::mutate(capture_process_date_time = as.POSIXct(paste(capture_process_date, capture_time, sep = " "), format = "%m/%d/%y %H:%M"),
                capture_process_date = as.POSIXct(capture_process_date, 
                                                  format = "%m/%d/%y"),
                # time with arbitrary date, need to figure out how to remove
                #capture_time = as.POSIXct(capture_time, format = "%H:%M"),
                #processing_time = as.POSIXct(processing_time, format = "%H:%M"),
                SVL_mm = as.numeric(SVL_mm),
                pre_tmt_mass_g = as.numeric(pre_tmt_mass_g),
                pre_mass_w_collar_Y_N = as.factor(pre_mass_w_collar_Y_N),
                post_tmt_mass_g = as.numeric(post_tmt_mass_g),
                post_mass_w_collar_Y_N = as.factor(post_mass_w_collar_Y_N),
                treatment_date_time = as.POSIXct(paste(treatment_date, processing_time, sep = " "), format = "%m/%d/%y %H:%M"),
                tmt_start_time = as.POSIXct(tmt_start_time, format = "%H:%M"),
                tmt_end_time = as.POSIXct(tmt_end_time, format = "%H:%M"),
                hematocrit_percent = as.numeric(hematocrit_percent),
                osmolality_mmol_kg = as.numeric(osmolality_mmol_kg),
                hemolyzed_Y_N = as.factor(hemolyzed_Y_N),
                cloacal_temp_C = as.numeric(cloacal_temp_C)
                ) %>%
  
  # add tmt group info
  left_join(read.csv("data/collared_identification.csv"),
            by = 'individual_ID') %>%
  
  # add osmolality values from blood samples
  #left_join(read.csv(osml)) %>%
  
  # add CEWL values from AquaFlux measurements
  #left_join(read.csv(CEWL)) %>%
  
                # format data classes
  dplyr::mutate(individual_ID = as.factor(individual_ID),
                sex_M_F = as.factor(sex),
                tmt = as.factor(tmt),
                radio_collar_mass_g = as.numeric(radio_collar_mass_g),
                
                #statistics
                capture_to_msmt = as.numeric(treatment_date_time - capture_process_date_time),
                tmt_time_elapsed = as.numeric(tmt_end_time - tmt_start_time),
                #test = capture_process_date == treatment_date,
                actual_pre_tmt_mass_g = pre_tmt_mass_g - radio_collar_mass_g*(pre_mass_w_collar_Y_N == "Y"),
                actual_post_tmt_mass_g = post_tmt_mass_g - radio_collar_mass_g*(post_mass_w_collar_Y_N == "Y"),
                tmt_mass_change_g = actual_post_tmt_mass_g - actual_pre_tmt_mass_g
                )

summary(lizards_collared)
```


## Widespread Lizards

The second dataset loaded in is for lizards we caught opportunistically and who we did not radio-collar or track throughout the summer.

```{r get data widespread lizards}
lizards_widespread <- read.csv("data/widespread_measurements.csv") %>%
  
                # format data classes
  dplyr::mutate(date = as.POSIXct(date, format = "%m/%d/%y"),
                # time with arbitrary date, need to figure out how to remove
                capture_time = as.POSIXct(capture_time, format = "%H:%M"),
                processing_time = as.POSIXct(processing_time, format = "%H:%M"),
                individual_ID = as.factor(individual_ID),
                SVL_mm = as.numeric(SVL_mm),
                mass_g = as.numeric(mass_g),
                sex_M_F = as.factor(sex_M_F),
                hematocrit_percent = as.numeric(hematocrit_percent),
                osmolality_mmol_kg = as.numeric(osmolality_mmol_kg),
                hemolyzed_Y_N = as.factor(hemolyzed.),
                cloacal_temp_C = as.numeric(cloacal_temp_C)
                ) %>%

  # add osmolality values from blood samples
  #left_join(read.csv(osml)) %>%
  
  # add CEWL values from AquaFlux measurements
  #left_join(read.csv(CEWL)) %>%

                #statistics
  dplyr::mutate(capture_to_msmt = as.numeric(processing_time - capture_time)
                )

summary(lizards_widespread)
```


## Variable Explanation



## Methods Stats

```{r capture to msmt}
# collared
mean(lizards_collared$capture_to_msmt, na.rm = TRUE)/60
median(lizards_collared$capture_to_msmt, na.rm = TRUE)/60
hist(lizards_collared$capture_to_msmt)
300/60

# widespread
mean(lizards_widespread$capture_to_msmt, na.rm = TRUE)
max(lizards_widespread$capture_to_msmt, na.rm = TRUE)
```

Most collared lizards were measured within 5 hours of capture, with the majority being measured within 2 hours of capture, and a few being measured the day after capture.

Widespread lizards were always measured within 5 hours of capture, and usually within 2 hours.

```{r tmt time}
mean(lizards_collared$tmt_time_elapsed, na.rm = T)
hist(lizards_collared$tmt_time_elapsed, na.rm = T)
```

Lizards were put in bins with shallow water for 1 hour, on average, to promote drinking behavior.


## Treatment Effects

Are there *ANY* differences between lizards who we gave water vs no water??

```{r tmt diffs}
lizards_collared %>%
  group_by(tmt, sex_M_F) %>%
  summarise(SVL_mean = mean(SVL_mm, na.rm = T),
            hct_mean = mean(hematocrit_percent, na.rm = T),
            osml_mean = mean(osmolality_mmol_kg, na.rm = T),
            mass_change_mean = mean(tmt_mass_change_g, na.rm = T)
            )
```

Potentially......

Are the widespread-caught lizards any different?

```{r wd diffs}
lizards_widespread %>%
  group_by(sex_M_F) %>%
  summarise(SVL_mean = mean(SVL_mm, na.rm = T),
            hct_mean = mean(hematocrit_percent, na.rm = T),
            osml_mean = mean(osmolality_mmol_kg, na.rm = T)
            )
```

Widespread lizards were simply the smaller ones that couldn't be radio-collared, so that's slightly different. Osmolality for those females seems higher than radio-collared control, but hct values seem on-par. 

We did recapture several widespread lizards, so I think it makes sense to put them on the same figures? Because that way, we can compare them to our control group. Maybe rename widespread -> "no intervention", control -> "sham tmt", and hydration -> "water tmt". :)




## Purge & Merge !

columns I want to have in BOTH:
- date
- ID
- SVL
- mass
- sex
- hct
- osml
- CEWL
- cloacal temp

```{r p n m}
widespread_clean <- lizards_widespread %>%
  dplyr::select(date,
                individual_ID,
                SVL_mm,
                mass_g,
                sex_M_F,
                hematocrit_percent,
                osmolality_mmol_kg,
                #CEWL,
                cloacal_temp_C
                ) %>%
  dplyr::mutate(tmt = as.factor("No Tmt"))

collared_clean <- lizards_collared %>%
  dplyr::select(date = capture_process_date,
                individual_ID,
                SVL_mm,
                mass_g = actual_pre_tmt_mass_g,
                sex_M_F,
                hematocrit_percent,
                osmolality_mmol_kg,
                #CEWL,
                cloacal_temp_C,
                tmt
                )

lizards_all <- rbind(widespread_clean, collared_clean) %>%
  dplyr::mutate(time_pt = as.factor(case_when(date == "2021-04-23 PDT" ~ 1, 
                                              date == "2021-04-24 PDT" ~ 1, 
                                              date == "2021-04-25 PDT" ~ 1,
                                              date == "2021-05-07 PDT" ~ 2,
                                              date == "2021-05-08 PDT" ~ 2,
                                              date == "2021-07-14 PDT" ~ 3)))

lizards_all$tmt <- factor(lizards_all$tmt,
                          levels = c("hydration", "control", "No Tmt"),
                          labels = c("Water Tmt", "Sham Tmt", "No Tmt")
                          )

lizards_all$time_pt <- factor(lizards_all$time_pt,
                              labels = c("April", "May", "July")
                              )

summary(lizards_all)
```






# Experimental Hydration Effects

```{r tmt mass change plot}
ggplot(data = lizards_collared) + 
  geom_boxplot(aes(x = tmt, 
                   y = tmt_mass_change_g, 
                   color = tmt
                   )) +
  geom_jitter(aes(x = tmt, 
                   y = tmt_mass_change_g, 
                   color = tmt
                   ), 
               size = 0.6,
               alpha = 0.6) +
  geom_point(aes(x = tmt, 
                   y = mean(tmt_mass_change_g, na.rm = T), 
                   color = tmt,
                   ), 
               size = 6,
               alpha = 1) +
  geom_hline(yintercept = 0) +
  theme_classic() + 
  xlab("Treatment Group") + 
  ylab("Mass Change (g)") + 
  #annotate("text", x = 1, y = 64, label = "a", size = 6) +
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```



```{r tmt mass change stats}
summary(lm(data = lizards_collared, tmt_mass_change_g ~ tmt))
```


So, it did result in a significant change in mass, but only 1.1g, on average.





# Tmt x Time Stats


```{r change by tmt * time}
# osmolality
summary(lm(data = lizards_all, osmolality_mmol_kg ~ tmt:time_pt))
TukeyHSD(aov(lm(data = lizards_all, osmolality_mmol_kg ~ tmt + time_pt)))

# hematocrit
summary(lm(data = lizards_all, hematocrit_percent ~ tmt:time_pt))

# mass
summary(lm(data = lizards_all, mass_g ~ tmt:time_pt))

# CEWL

```

There was NO effect of treatment on osmolality, hematocrit, or mass change throughout the season. 

So, we can carry on with all 3 treatment groups aggregated.


# Time Stats

```{r change over time only}
# only do for individuals measured at time points 1 and 2
indiv_over_time <- lizards_all %>%
  group_by(individual_ID) %>%
  summarise(n = n()) %>%
  dplyr::filter(n > 1)
lizards_over_time <- lizards_all %>%
  dplyr::filter(individual_ID %in% indiv_over_time$individual_ID)

# now the following stats will be "for lizards measured at >1 time pt..."

# osmolality
summary(lmerTest::lmer(data = lizards_over_time, 
                       osmolality_mmol_kg ~ time_pt + (1|individual_ID)))

# hematocrit
summary(lmerTest::lmer(data = lizards_over_time, 
                       hematocrit_percent ~ time_pt + (1|individual_ID)))

# mass
summary(lmerTest::lmer(data = lizards_over_time, 
                       mass_g ~ time_pt + (1|individual_ID)))

# CEWL

```

On average, with individual variation accounted for, osmolality and hematocrit both increased from April to May (but not for July). Mass increased from April to May, then decreased in July. 



# Figures

## Osmolality

```{r osml ~ time fig}
ggplot(data = lizards_over_time) + 
  geom_point(aes(x = time_pt, 
                   y = osmolality_mmol_kg, 
                group = individual_ID,
                color = individual_ID), 
               size = 2,
               alpha = 1) +
  geom_line(aes(x = time_pt, 
                   y = osmolality_mmol_kg, 
                group = individual_ID,
                color = individual_ID),
            alpha = 0.8) +
  theme_classic() + 
  xlab("Time (not to scale)") + 
  ylab("Osmolality (mmol/kg)") + 
  xlim(c("April", "May")) +
  #annotate("text", x = 1, y = 64, label = "a", size = 6) +
  #scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```


## Hct

```{r hct ~ time fig}
ggplot(data = lizards_over_time) + 
  geom_point(aes(x = time_pt, 
                   y = hematocrit_percent, 
                group = individual_ID,
                color = individual_ID), 
               size = 2,
               alpha = 1) +
  geom_line(aes(x = time_pt, 
                   y = hematocrit_percent, 
                group = individual_ID,
                color = individual_ID),
            alpha = 0.8) +
  theme_classic() + 
  xlab("Time (not to scale)") + 
  ylab("Hematocrit (%)") + 
  xlim(c("April", "May")) +
  #annotate("text", x = 1, y = 64, label = "a", size = 6) +
  #scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```


It might make sense to remove what looks like a few outliers from April. If those lizards had very little blood drawn, their hct values would be unreliable.



## Mass

```{r mass ~ time fig}
ggplot(data = lizards_over_time) + 
  geom_point(aes(x = time_pt, 
                   y = mass_g, 
                group = individual_ID,
                color = individual_ID), 
               size = 2,
               alpha = 1) +
  geom_line(aes(x = time_pt, 
                   y = mass_g, 
                group = individual_ID,
                color = individual_ID),
            alpha = 0.8) +
  theme_classic() + 
  xlab("Time (not to scale)") + 
  ylab("Mass (g)") + 
  #annotate("text", x = 1, y = 64, label = "a", size = 6) +
  #scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```




# CEWL

## Time

## Body Temp

## Body Condition





































