---
title: "Blunt-nosed Leopard Lizard Data Analysis: Water Physiology Nuances"
author: "Savannah Weaver"
output: 
  rmdformats::html_clean:
    highlight: tango
    thumbnails: FALSE
    toc: TRUE
    toc_depth: 3
---


# Packages

```{r setup, echo=T, results='hide', message=FALSE}
if (!require("tidyverse")) install.packages("tidyverse") 
library("tidyverse")# for dplyr workflow of calculations
if (!require("UsingR")) install.packages("UsingR") 
library("UsingR")
if (!require("lme4")) install.packages("lme4") 
library("lme4")
if (!require("lmerTest")) install.packages("lmerTest") 
library("lmerTest")
if (!require("broom")) install.packages("broom")
library("broom") # lmer model export
if (!require("broom.mixed")) install.packages("broom.mixed")
library("broom.mixed") # lmer model export
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
library("RColorBrewer")
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # for multi-ggplot figs
if (!require("Hmisc")) install.packages("Hmisc")
library("Hmisc") # correlation matrix
if (!require("emmeans")) install.packages("emmeans")
library("emmeans") # marginal means & confints
```





# Set Colors & Themes

```{r}
savs_ggplot_theme <- theme_bw() + 
                     theme(text = element_text(color = "black", 
                                               family = "sans", 
                                               size = 12),
                           axis.text = element_text(color = "black", 
                                                    family = "sans", 
                                                    size = 8),
                           legend.text = element_text(color = "black", 
                                                      family = "sans", 
                                                      size = 8),
                           legend.text.align = 0,
                           legend.position = "bottom",
                           #plot.margin = margin(t = 6, r = 6, b = 0, l = 1, unit = "pt")
                           ) #+
  
  #theme_classic() + 
  #guides(shape = guide_legend(nrow = 2, byrow = TRUE))
  
savs_ggplot_theme_corrs <- theme_classic() +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 20),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 18),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 18),
        legend.text.align = 0,
        legend.position = "right"
        )
  
# custom colors
Male <- brewer.pal(11, "RdYlBu")[c(11)]
Fem <- brewer.pal(11, "RdYlBu")[c(2)]
my_colors <- c(Fem, Male)
my_shapes <- c(19,15)
my_shapes_open <- c(21,22)

# corr colors
n3 <- brewer.pal(9, "YlGnBu")[c(4, 6, 8)]
n2 <- brewer.pal(9, "YlGnBu")[c(6, 8)]
n1 <- brewer.pal(9, "YlGnBu")[c(6)]


# labels for plots for repeat measures lizards only
my_labels <- c("April\nF\n(n=11)", "April\nM\n(n=22)",
               "May\nF\n(n=10)", "May\nM\n(n=17)",
               "July\nF\n(n=2)", "July\nM\n(n=12)")
# labels for plots for full dataset of all lizards
my_labels_full <- c("April\nF\n(n=26)", "April\nM\n(n=41)",
                     "May\nF\n(n=15)", "May\nM\n(n=22)",
                     "July\nF\n(n=3)", "July\nM\n(n=12)")
```










# Background 



# Load in Data

This data was collected during fieldwork with Blunt-nosed Leopard Lizards (*Gambelia sila*) in the Elkhorn Plain of the Carrizo Plain National Monument in California, USA during the 2021 active season (April-August). See published paper (doi) for complete methods and discussion of these analyses. 


```{r}
dat <- read_rds("./data/G_sila_clean_full_dat.RDS") %>%
  mutate(month_sex = factor(paste(month, sex_M_F),
                            levels = c("April Female", "April Male",
                                       "May Female", "May Male",
                                       "July Female", "July Male")),
         date_chunk_pretty = factor(case_when(month == "April" ~ "Apr 23-25",
                                              month == "May" ~ "May 7-9",
                                              month == "July" ~ "Jul 14"),
                                    levels = c("Apr 23-25", "May 7-9", "Jul 14")),
         gravid_Y_N = factor(gravid_Y_N, levels = c("Not Gravid", "Gravid")))
summary(dat)
unique(dat$radio_collar_serial)

# summary statistics
dat %>%
  summarise(mean(CEWL_g_m2h, na.rm = T), sd(CEWL_g_m2h, na.rm = T),
            mean(osmolality_mmol_kg_mean, na.rm = T), 
            sd(osmolality_mmol_kg_mean, na.rm = T),
            mean(hematocrit_percent, na.rm = T), sd(hematocrit_percent, na.rm = T),
            mean(mass_g), sd(mass_g),
            mean(SVL_mm), sd(SVL_mm),
            mean(SMI), sd(SMI))
```




## Variable Explanation






## Repeat Subset

I want to make some sub-dataframes for when I want to do stats for only a subset of individuals.

First, get a dataframe of only individuals that had more than one measurement taken, and note how many individuals had x measurements:

```{r}
# how many lizards were measured 1, 2, or 3 times?
note_repeats <- dat %>%
  group_by(individual_ID) %>%
  summarise(n = n()) %>%
  group_by(n) %>%
  summarise(n_lizards = n())
note_repeats 

# this is how many individual lizards we measured :)
sum(note_repeats$n_lizards) 

# get subsetted data
dat_repeats <- dat %>%
  group_by(individual_ID) %>%
  mutate(n = n()) %>%
  dplyr::filter(n>1)

# check these match 'note_repeats'
dat_repeats %>%
  group_by(individual_ID) %>%
  summarise(n = n()) %>%
  group_by(n) %>%
  summarise(n_lizards = n())
```



## Female Subset

To look at reproductive effort ~ water balance, I will make another subset for only the females we collected gravidity/egg data on:

```{r}
dat_repro_females <- dat %>%
  dplyr::filter(sex_M_F == "Female") %>%
  dplyr::filter(complete.cases(gravid_Y_N)) %>%
  group_by(individual_ID) %>%
  mutate(n = n(),
         n = factor(n)) 
summary(dat_repro_females)
```


It's good to note that I measured reproductive effort at **two time points** for 9 females, and only at one time point for 10 females.

Check whether the single-measurement females are all from the first weekend or not:

```{r}
table(dat_repro_females$week, dat_repro_females$n)
dat_repro_females %>%
  dplyr::filter(n == 1) %>%
  summary() # 4 g, 6 ng
dat_repro_females %>%
  dplyr::filter(n == 2) %>%
  dplyr::filter(week == 3) %>%
  summary() # 8 g, 1 ng
table(dat_repro_females$gravid_Y_N, dat_repro_females$n)
```


For the females who were only measured once, 8 of them are from the first weekend and 2 of them are from the third weekend.

Also save a subset of the females that were measured twice:


```{r}
dat_repeat_females <- dat_repro_females %>%
  dplyr::filter(n == 2)
```




# Does SVL change over time?

I want to know whether SVL changed over time for the individuals who had more than one measurement taken. If it does, then body condition is certainly more informative than mass.

```{r}
# model
summary(lmerTest::lmer(data = dat_repeats, 
                   SVL_mm ~ week_num + (1|individual_ID)))
```


There is no meaningful change in SVL for the individuals we were able to measure, so I could use either mass or body condition to estimate body size.



# Check CEWL Covariates

How much is CEWL impacted by the body temperature, ambient temperature, and ambient VPD at the time of CEWL measurement? I will investigate this using *all* measurements.

It's possible that I may find a relationship between ambient temp and CEWL but not body temp and CEWL because the body temp measurements we took were unfortunately only to the nearest degree...


## Plot

First, make a simple plot of each to look for relationships visually.

```{r}
ggplot(dat) +
  aes(x = cloacal_temp_C,
      y = CEWL_g_m2h,
      color = week
      #color = sex_M_F
        ) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw()
ggplot(dat) +
  aes(x = msmt_temp_C,
      y = CEWL_g_m2h,
      color = week
      #color = sex_M_F
        ) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw()
ggplot(dat) +
  aes(x = msmt_VPD_kPa,
      y = CEWL_g_m2h,
      color = week
      #color = sex_M_F
        ) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw()
```


Across dates, CEWL had no relationship with body temp, ambient temp, or ambient VPD at the time of capture. However, when separated by date, there were relationships. CEWL had a positive relationship with body temp, ambient temp, and ambient VPD, but only for week 3 measurements. There were no relationships when separated by treatment group. 


## Models

Now, check linear models of the effect of each of body temp, ambient temp, and ambient VPD with the interaction of date on CEWL.

```{r}
summary(lm(data = dat,
           CEWL_g_m2h ~ cloacal_temp_C*week))
summary(lm(data = dat,
           CEWL_g_m2h ~ msmt_temp_C*week))
summary(lm(data = dat,
           CEWL_g_m2h ~ msmt_VPD_kPa*week))
```

CEWL had a positive relationship with each of body temp, ambient temp, and ambient VPD, *but only for the measurements taken in May* (week 3).


## Variance

Let's also check the distribution/range of these covariates, because based on the figures, week 3 had a pretty small range of temperatures and VPDs.

```{r}
dat %>%
  group_by(week) %>%
  summarise(ctemp_range = max(cloacal_temp_C, na.rm = T) - 
              min(cloacal_temp_C, na.rm = T),
            mtemp_range = max(msmt_temp_C, na.rm = T) - 
              min(msmt_temp_C, na.rm = T),
            mVPD_range = max(msmt_VPD_kPa, na.rm = T) - 
              min(msmt_VPD_kPa, na.rm = T))
```

Well... the ranges for week 3 are not small enough to be meaningless. So, we will have to keep in mind that temperature and VPD should be retained in CEWL models as a covariate.




# Water Balance Metrics

What is the interrelation among water balance metrics CEWL, osmolality, hematocrit, and mass/SMI. I'll use all measurements for this. *I'm not sure if we will include hematocrit in the final results, though.*

## Models

Check a simple linear model between each set of paired measurements:

```{r}
# CEWL ~
anova(lm(data = dat,
           CEWL_g_m2h ~ osmolality_mmol_kg_mean*month))
summary(lm(data = dat,
           CEWL_g_m2h ~ osmolality_mmol_kg_mean*month*sex_M_F))
anova(lm(data = dat,
           CEWL_g_m2h ~ hematocrit_percent))
anova(lm(data = dat,
           CEWL_g_m2h ~ mass_g))
anova(lm(data = dat,
           CEWL_g_m2h ~ SMI))

# osmolality ~
anova(lm(data = dat,
           osmolality_mmol_kg_mean ~ hematocrit_percent))
anova(lm(data = dat,
           osmolality_mmol_kg_mean ~ mass_g))
anova(lm(data = dat,
           osmolality_mmol_kg_mean ~ SMI))

# hct
anova(lm(data = dat,
           hematocrit_percent ~ SMI*sex_M_F))
```

Notes on these relationships:

- CEWL ~ osml had significant interaction of osml* month (osml* mo* sex was not sig)
- CEWL ~ hct was sig but had no extra interactions (plot direct relationship only)
- CEWL ~ SMI was sig (mass was not; & SMI did not interact with mo or sex)
- osml & hct corr, but with no interactions of mo or sex
- osml also corr with SMI but not mass (& no interactions with mo or sex)
**^all of the above based on anova SS, so assessing whether these things explain variation in response (CEWL/osml)**

For MS:
Plasma osmolality explained a significant amount of the variation in CEWL (SLR: F(1,110) = 5.07, p = 0.03), with a marginally nonsignificant interaction with measurement time period (SLR: F(2,110) = 2.77, p = 0.07). CEWL could also be explained by hematocrit (SLR: F(1,114) = 9.86, p = 0.002), or body condition (SLR: F(1,116) = 3.89, p = 0.05), but not lizard mass (SLR: F(1,116) = 1.44, p = 0.2). In turn, variation in plasma osmolality could be explained by hematocrit (SLR: F(1,114) = 15.07, p < 0.001) or body condition (SLR: F(1,114) = 5.30, p = 0.02), but not lizard mass (SLR: F(1,114) < 0.001, p = 1). Hematocrit and body condition were also significantly related (SLR: F(1,112) = 9.03, p = 0.003), with an interaction of lizard sex (SLR: F(1,112) = 7.97, p = 0.005). 

Plots to make *nicely*:

- CEWL ~ osml*mo
  - Fig caption stats: adj-R-sq = 0.16
- CEWL ~ hct g
- CEWL ~ SMI (& note about mass) g
- osml ~ hct g
- osml ~ SMI g
- hct ~ SMI*sex



## Plot CEWL ~ Osmolality


```{r CEWL osml fig}
ggplot(dat) + 
  aes(x = osmolality_mmol_kg_mean,
      y = CEWL_g_m2h,
      color = date_chunk_pretty
      #color = sex_M_F
      ) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1) + 
  scale_color_manual(values = n3, name = "Measurement\nPeriod") + 
  xlab("Plasma Osmolality (mmol/kg)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim(300, 450) +
  #annotate(geom = "text", x = 430, y = 24, label = bquote(''*R^2*'=0.033'), 
   #        size = 7) +
  #ylim(0, 25) +
  savs_ggplot_theme_corrs -> CEWL_osml_fig
CEWL_osml_fig

# export figure
#ggsave(filename = "CEWL_osml_fig.jpeg",
 #      plot = CEWL_osml_fig,
  #     path = "./results_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 6, height = 4)
```


## Plot CEWL ~ SMI


```{r CEWL SMI fig}
ggplot(dat) + 
  aes(x = SMI,
      y = CEWL_g_m2h,
      #color = tmt
      #color = month_sex
      ) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1,
              color = n1) + 
  xlab("Body Condition (g')") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim(300, 450) +
  #annotate(geom = "text", x = 430, y = 24, label = bquote(''*R^2*'=0.033'), 
   #        size = 7) +
  #ylim(0, 25) +
  savs_ggplot_theme_corrs -> CEWL_SMI_fig
CEWL_SMI_fig
```



## Plot CEWL ~ Hematocrit


```{r CEWL hct fig}
ggplot(dat) + 
  aes(x = hematocrit_percent,
      y = CEWL_g_m2h,
      #color = tmt
      #color = month_sex
      ) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1,
              color = n1) + 
  xlab("Hematocrit (%)") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  #ylab("") +
  #xlim(300, 450) +
  #annotate(geom = "text", x = 430, y = 24, label = bquote(''*R^2*'=0.033'), 
   #        size = 7) +
  #ylim(0, 25) +
  savs_ggplot_theme_corrs -> CEWL_hct_fig
CEWL_hct_fig
```








## Plot Osmol ~ SMI


```{r osml SMI fig}
ggplot(dat) + 
  aes(x = SMI,
      y = osmolality_mmol_kg_mean,
      #color = tmt
      #color = month_sex
      ) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1,
              color = n1) + 
  xlab("Body Condition (g')") + 
  ylab("Osmolality (mmol/kg)") + 
  #ylab("") +
  #xlim(300, 450) +
  #annotate(geom = "text", x = 430, y = 24, label = bquote(''*R^2*'=0.033'), 
   #        size = 7) +
  #ylim(0, 25) +
  savs_ggplot_theme_corrs -> osml_SMI_fig
osml_SMI_fig
```






## Plot Osmol ~ Hematocrit


```{r osml hct fig}
ggplot(dat) + 
  aes(x = hematocrit_percent,
      y = osmolality_mmol_kg_mean,
      #color = tmt
      #color = month_sex
      ) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1,
              color = n1) + 
  xlab("Hematocrit (%)") + 
  ylab("Osmolality (mmol/kg)") + 
  #ylab("") +
  #xlim(300, 450) +
  #annotate(geom = "text", x = 430, y = 24, label = bquote(''*R^2*'=0.033'), 
   #        size = 7) +
  #ylim(0, 25) +
  scale_color_manual(values = n1) + 
  savs_ggplot_theme_corrs -> osml_hct_fig
osml_hct_fig
```







## Plot Hct ~ SMI


```{r hct SMI fig}
ggplot(dat) + 
  aes(x = SMI,
      y = hematocrit_percent,
      color = sex_M_F,
      #color = month_sex
      ) +
  geom_point(size = 1, 
             alpha = 0.4) + 
  stat_smooth(formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1) + 
  xlab("Body Condition (g')") + 
  ylab("Hematocrit (%)") + 
  #ylab("") +
  #xlim(300, 450) +
  #annotate(geom = "text", x = 430, y = 24, label = bquote(''*R^2*'=0.033'), 
   #        size = 7) +
  #ylim(0, 25) +
  scale_color_manual(values = n2) +
  savs_ggplot_theme_corrs -> hct_SMI_fig
hct_SMI_fig
```





## Export Multi-Plot


```{r}
ggarrange(CEWL_SMI_fig, CEWL_hct_fig,
          osml_SMI_fig, osml_hct_fig,
          ncol = 2, nrow = 2,
          #labels = c("A", "B", "C"),
          common.legend = TRUE,
          legend = "bottom"
          ) -> water_balance_interrelation
#water_balance_interrelation

ggsave(filename = "water_balance_interrelation.pdf",
       plot = water_balance_interrelation,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       width = 200, height = 150
       )
```













## Week Interaction

Redo CEWL ~ osmolality model with an interaction of date:

```{r}
summary(lm(data = dat, CEWL_g_m2h ~ osmolality_mmol_kg_mean*week))
```


Once again, the relationship is only significant for week 3...\



Also look at for hematocrit:

```{r}
summary(lm(data = dat,
           CEWL_g_m2h ~ hematocrit_percent*week))
summary(lm(data = dat,
           osmolality_mmol_kg_mean ~ hematocrit_percent*week))
```

The relationships between CEWL and osmolality with hematocrit are not different across weeks.



## Lag Effect?

What if I tested week 3 CEWL ~ week 1 osml, or visa versa? This is to look for a lag effect on their relationship, although the time period is likely too long.

First, organize the data:

```{r}
w3_CEWL <- dat %>%
  dplyr::filter(week == 3) %>%
  dplyr::select(individual_ID, CEWL_g_m2h)
w1_osml <- dat %>%
  dplyr::filter(week == 1) %>%
  dplyr::select(individual_ID, osmolality_mmol_kg_mean)
w3_vs_w1 <- w1_osml %>%
  left_join(w3_CEWL, by = 'individual_ID') %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean, CEWL_g_m2h))
w1_CEWL <- dat %>%
  dplyr::filter(week == 1) %>%
  dplyr::select(individual_ID, CEWL_g_m2h)
w3_osml <- dat %>%
  dplyr::filter(week == 3) %>%
  dplyr::select(individual_ID, osmolality_mmol_kg_mean) 
w1_vs_w3 <- w3_osml %>%
  left_join(w1_CEWL, by = 'individual_ID') %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean, CEWL_g_m2h))
```


Run models:

```{r}
summary(lm(data = w3_vs_w1, CEWL_g_m2h ~ osmolality_mmol_kg_mean))
summary(lm(data = w1_vs_w3, CEWL_g_m2h ~ osmolality_mmol_kg_mean))
```

It seems very unlikely there is a lag effect, at least at this time scale. t and F stats were both nonsig for each comparisons.



# Supplemental Hydration Treatment Effects


## Calculate Deltas

First, calculate the change in CEWL, osmolality, and mass for each treatment individual that was recaptured between April and May.

```{r}
starting_values <- dat_repeats %>%
  # look only at lizards in hydration or sham tmt
  dplyr::filter(tmt != "No Tmt") %>%
  # get the values for April measurements only 
  dplyr::filter(month == "April") %>%
  dplyr::select(individual_ID,
                CEWL_April = CEWL_g_m2h,
                osml_April = osmolality_mmol_kg_mean,
                mass_April = mass_g)
dat_tmt_change <- dat_repeats %>%
  # look only at lizards in hydration or sham tmt
  dplyr::filter(tmt != "No Tmt") %>%
  # add starting values
  left_join(starting_values, by = 'individual_ID') %>%
  # make sure we have complete data
  dplyr::filter(complete.cases(mass_April, CEWL_April)) %>%
  # calculate difference for each msmt versus in April
  mutate(mass_change_from_April = mass_g - mass_April,
         CEWL_change_from_April = CEWL_g_m2h - CEWL_April,
         osml_change_from_April = osmolality_mmol_kg_mean - osml_April)
dat_tmt_change_May_only <- dat_tmt_change %>%
  dplyr::filter(month == "May")
```



## Plot Change by Tmt

### Mass

```{r}
ggplot(dat_tmt_change) +
  aes(x = month,
      y = mass_change_from_April,
      group = individual_ID,
      color = tmt
      ) +
  xlab("Measurement Month") + 
  ylab(expression(Delta ~ 'Lizard Mass (g)')) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(values = my_colors, name = "Sex") +
  savs_ggplot_theme -> tmt_delta_mass
tmt_delta_mass
```


### Osmolality

```{r}
ggplot(dat_tmt_change) +
  aes(x = month,
      y = osml_change_from_April,
      group = individual_ID,
      color = tmt
      ) +
  xlab("Measurement Month") + 
  ylab(expression(Delta ~ 'Osmolality (mmol '*kg^-1*')')) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(values = my_colors, name = "Sex") +
  savs_ggplot_theme -> tmt_delta_osml
tmt_delta_osml
```




### CEWL

```{r}
ggplot(dat_tmt_change) +
  aes(x = month,
      y = CEWL_change_from_April,
      group = individual_ID,
      color = tmt
      ) +
  xlab("Measurement Month") + 
  ylab(expression(Delta ~ 'CEWL (g '*m^-2*' '*h^-1*')')) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(values = my_colors, name = "Sex") +
  savs_ggplot_theme -> tmt_delta_CEWL
tmt_delta_CEWL
```


It does *not* look like treatment led to any differences in the amount or direction of change in lizard mass, CEWL, or osmolality. Changes were also not consistent by sex.

### Export Multi-Panel Fig

Save plots:

```{r}
ggarrange(tmt_delta_CEWL,
          tmt_delta_osml, 
          tmt_delta_mass,
          ncol = 1, nrow = 3,
          labels = c("A", "B", "C"),
          common.legend = TRUE,
          legend = "bottom"
          ) -> tmt_delta_multi_fig

ggsave(filename = "tmt_delta_multi_fig.pdf",
       plot = tmt_delta_multi_fig,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       width = 80, height = 210)
```



## Stat Tests of Change


What about mass change *during* treatment, when we either gave lizards water to drink or a sham treatment. 

```{r tmt mass change stats}
summary(lm(data = dat_tmt_change, tmt_mass_change_g ~ tmt))
emmeans(lm(data = dat_tmt_change, tmt_mass_change_g ~ tmt), "tmt")
```

So, it did result in a significant change in mass within the same-day water supplementation tmt (vs sham), but only ~1.4g difference, on average.





These lm's are doing the same thing as a two-sample t-test, checking whether the means are statistically different. 

```{r}
summary(lm(data = dat_tmt_change_May_only, mass_change_from_April ~ tmt))
summary(lm(data = dat_tmt_change_May_only, CEWL_change_from_April ~ tmt))
summary(lm(data = dat_tmt_change_May_only, osml_change_from_April ~ tmt))
# stats in paper are from these 3 models
summary(lm(data = dat_tmt_change, mass_change_from_April ~ tmt))
summary(lm(data = dat_tmt_change, CEWL_change_from_April ~ tmt))
summary(lm(data = dat_tmt_change, osml_change_from_April ~ tmt))
```

Change was not significantly different between treatments for any of the variables. 


Was change overall (across tmts) different from zero?

```{r}
t.test(dat_tmt_change_May_only$mass_change_from_April)
t.test(dat_tmt_change_May_only$CEWL_change_from_April)
t.test(dat_tmt_change_May_only$osml_change_from_April)
```

On average, lizards who were measured in both April and May increased mass, but did not experience a change in osmolality or CEWL. 





# Change Over Time


## Plot Delta ~ Time*Sex

### Mass

```{r}
ggplot(dat_tmt_change) +
  aes(x = month,
      y = mass_change_from_April,
      group = individual_ID,
      color = sex_M_F,
      shape = sex_M_F
      ) +
  xlab("Measurement Month") + 
  ylab(expression(Delta ~ 'Lizard Mass (g)')) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes, name = "Sex") +
  savs_ggplot_theme -> delta_mass_sex
delta_mass_sex
```


### Osmolality

```{r}
ggplot(dat_tmt_change) +
  aes(x = month,
      y = osml_change_from_April,
      group = individual_ID,
      color = sex_M_F,
      shape = sex_M_F
      ) +
  xlab("Measurement Month") + 
  ylab(expression(Delta ~ 'Osmolality (mmol '*kg^-1*')')) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes, name = "Sex") +
  savs_ggplot_theme -> delta_osml_sex
delta_osml_sex
```




### CEWL

```{r}
ggplot(dat_tmt_change) +
  aes(x = month,
      y = CEWL_change_from_April,
      group = individual_ID,
      color = sex_M_F,
      shape = sex_M_F
      ) +
  xlab("Measurement Month") + 
  ylab(expression(Delta ~ 'CEWL (g '*m^-2*' '*h^-1*')')) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes, name = "Sex") +
  savs_ggplot_theme -> delta_CEWL_sex
delta_CEWL_sex
```



### Export Multi-Panel Fig

Save plots:

```{r}
ggarrange(delta_CEWL_sex,
          delta_osml_sex, 
          delta_mass_sex,
          ncol = 3, nrow = 1,
          #ncol = 1, nrow = 3,
          labels = c("A", "B", "C"),
          common.legend = TRUE,
          legend = "bottom"
          ) -> delta_multi_fig_sex

ggsave(filename = "delta_multi_fig_sex.pdf",
       plot = delta_multi_fig_sex,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       width = 200, height = 80
       #width = 80, height = 210
       )
```



## Time*Sex Marginal Means

I want to know whether the average values for all the lizards we measured changes over time. To test this, I can run a linear model, then use emmeans to get the marginal means and confidence intervals for each time point. 

**Do this only for lizards with repeat measurements.**

info for using emmeans with interactions:
https://cran.r-project.org/web/packages/emmeans/vignettes/interactions.html


Check n's:

```{r}
dat_repeats %>%
  group_by(month, sex_M_F) %>%
  summarise(n = n())
```




### Mass

```{r}
time_sex_diffs_mass <- lmerTest::lmer(data = dat_repeats, mass_g ~ month*sex_M_F +
                                        (1|individual_ID))
anova(time_sex_diffs_mass, type = "1", ddf = "Kenward-Roger")

# mean values
emmeans_mass <- data.frame(emmeans(time_sex_diffs_mass, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "Mass")

# stat diffs by month
time_emmeans_mass <- data.frame(emmeans(time_sex_diffs_mass, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "Mass")

# stat diffs by sex
sex_emmeans_mass <- data.frame(emmeans(time_sex_diffs_mass, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "Mass")
```


### Body Condition

```{r}
time_sex_diffs_SMI <- lmerTest::lmer(data = dat_repeats, SMI ~ month*sex_M_F +
                                        (1|individual_ID))
anova(time_sex_diffs_SMI, type = "1", ddf = "Kenward-Roger")

# mean values
emmeans_SMI <- data.frame(emmeans(time_sex_diffs_SMI, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "SMI")

# stat diffs by month
time_emmeans_SMI <- data.frame(emmeans(time_sex_diffs_SMI, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "SMI")

# stat diffs by sex
sex_emmeans_SMI <- data.frame(emmeans(time_sex_diffs_SMI, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "SMI")
```




### Osmolality

*Remember, for some reason VPD was very significant when added to the osmolality lm...*

```{r}
time_sex_diffs_osml <- lmerTest::lmer(data = dat_repeats, 
                                      osmolality_mmol_kg_mean ~ month*sex_M_F +
                                        (1|individual_ID))
anova(time_sex_diffs_osml, type = "1", ddf = "Kenward-Roger")

# mean values
emmeans_osml <- data.frame(emmeans(time_sex_diffs_osml, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "Osmolality")

# stat diffs by month
time_emmeans_osml <- data.frame(emmeans(time_sex_diffs_osml, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "Osmolality")

# stat diffs by sex
sex_emmeans_osml <- data.frame(emmeans(time_sex_diffs_osml, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "Osmolality")
```




### CEWL

```{r}
time_sex_diffs_CEWL <- lmerTest::lmer(data = dat_repeats, 
                                      CEWL_g_m2h ~ month*sex_M_F +
                                        (1|individual_ID))
anova(time_sex_diffs_CEWL, type = "1", ddf = "Kenward-Roger")

# check covariates
time_sex_diffs_CEWL_check_covaries <- lmerTest::lmer(data = dat_repeats, 
                                                     CEWL_g_m2h ~ month*sex_M_F
                                         + msmt_temp_C + msmt_VPD_kPa +
                                        (1|individual_ID))
anova(time_sex_diffs_CEWL_check_covaries, type = "1", ddf = "Kenward-Roger") # not important at all
59/0.09 # 655.5
54/0.14 # 385.7

# mean values
emmeans_CEWL <- data.frame(emmeans(time_sex_diffs_CEWL, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "CEWL")

# stat diffs by month
time_emmeans_CEWL <- data.frame(emmeans(time_sex_diffs_CEWL, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "CEWL")

# stat diffs by sex
sex_emmeans_CEWL <- data.frame(emmeans(time_sex_diffs_CEWL, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "CEWL")
```



### Check Assumptions

```{r}
plot(time_sex_diffs_CEWL)
plot(time_sex_diffs_osml)
plot(time_sex_diffs_mass)
```



### Merge dfs

```{r}
emmeans_all <- emmeans_CEWL %>%
  rbind(emmeans_osml) %>%
  rbind(emmeans_mass) %>%
  rbind(emmeans_SMI) %>%
  mutate(month_sex = factor(paste(month, sex_M_F),
                            levels = c("April Female", "April Male",
                                       "May Female", "May Male",
                                       "July Female", "July Male")))

# all pairwise stats together
emmean_time_stats <- time_emmeans_CEWL %>%
  rbind(time_emmeans_osml) %>%
  rbind(time_emmeans_mass) %>%
  rbind(time_emmeans_SMI)
emmean_sex_stats <- sex_emmeans_CEWL %>%
  rbind(sex_emmeans_osml) %>%
  rbind(sex_emmeans_mass) %>%
  rbind(sex_emmeans_SMI)
```







## Plot MMeans Time*Sex

### Mass

```{r}
ggplot() +
  geom_jitter(data = dat_repeats, # dat or dat_repeats ?
             aes(x = month_sex,
                 y = mass_g,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
              position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = emmeans_all[emmeans_all$measurement == "Mass",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = emmeans_all[emmeans_all$measurement == "Mass",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                 color = sex_M_F),
                width = .4) +
  xlab("") + 
  ylab("Lizard Mass (g)") +
  # sex differences = NONE
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 61, label = "A", size = 3) +
  annotate(geom = "text", x = 4, y = 61, label = "A", size = 3) +
  annotate(geom = "text", x = 6, y = 61, label = "A", size = 3) +
  # time differences for FEMALES
  annotate(geom = "text", x = 1, y = 61, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 61, label = "b", size = 3) +
  annotate(geom = "text", x = 5, y = 61, label = "c", size = 3) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels) +
  ylim(17, 61) +
  savs_ggplot_theme -> mass_emmean_plot
mass_emmean_plot
```



### Body Condition

```{r}
ggplot() +
  geom_jitter(data = dat_repeats, # dat or dat_repeats ?
             aes(x = month_sex,
                 y = SMI,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
              position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = emmeans_all[emmeans_all$measurement == "SMI",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = emmeans_all[emmeans_all$measurement == "SMI",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                 color = sex_M_F),
                width = .4) +
  xlab("") + 
  ylab("Body Condition (g')") +
  # sex differences
  annotate(geom = "text", x = 3.5, y = 45, label = "*", size = 6) +
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 54, label = "AB", size = 3) +
  annotate(geom = "text", x = 4, y = 54, label = "A", size = 3) +
  annotate(geom = "text", x = 6, y = 54, label = "B", size = 3) +
  # time differences for FEMALES
  annotate(geom = "text", x = 1, y = 54, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 54, label = "b", size = 3) +
  annotate(geom = "text", x = 5, y = 54, label = "a", size = 3) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels) +
  ylim(20, 55) +
  savs_ggplot_theme -> SMI_emmean_plot
SMI_emmean_plot
```





### Osmolality

```{r}
ggplot() +
  geom_jitter(data = dat_repeats,  # dat or dat_repeats ?
             aes(x = month_sex,
                 y = osmolality_mmol_kg_mean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
              position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = emmeans_all[emmeans_all$measurement == "Osmolality",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = emmeans_all[emmeans_all$measurement == "Osmolality",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                 color = sex_M_F),
                width = .4) +
  xlab("") + 
  ylab(bquote('Osmolality (mmol '*kg^-1*')')) +
  # sex differences
  annotate(geom = "text", x = 1.5, y = 416, label = "*", size = 6) +
  annotate(geom = "text", x = 3.5, y = 416, label = "*", size = 6) +
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 430, label = "A", size = 3) +
  annotate(geom = "text", x = 4, y = 430, label = "A", size = 3) +
  annotate(geom = "text", x = 6, y = 430, label = "B", size = 3) +
  # time differences for FEMALES
  annotate(geom = "text", x = 1, y = 430, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 430, label = "a", size = 3) +
  annotate(geom = "text", x = 5, y = 430, label = "a", size = 3) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels) +
  scale_y_continuous(limits = c(320, 440), breaks = c(320, 360, 400, 440)) +
  savs_ggplot_theme -> osml_emmean_plot
osml_emmean_plot
```




### CEWL

```{r}
ggplot() +
  geom_jitter(data = dat_repeats, # dat or dat_repeats ?
             aes(x = month_sex,
                 y = CEWL_g_m2h,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
             position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = emmeans_all[emmeans_all$measurement == "CEWL",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = emmeans_all[emmeans_all$measurement == "CEWL",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                    color = sex_M_F),
                width = .4) +
  # sex differences
  annotate(geom = "text", x = 3.5, y = 16, label = "*", size = 6) +
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 19, label = "A", size = 3) +
  annotate(geom = "text", x = 4, y = 19, label = "B", size = 3) +
  annotate(geom = "text", x = 6, y = 19, label = "A", size = 3) +
  # time differences for FEMALES
  annotate(geom = "text", x = 1, y = 19, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 19, label = "a", size = 3) +
  annotate(geom = "text", x = 5, y = 19, label = "a", size = 3) +
  xlab("") + 
  ylab(bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels) +
  savs_ggplot_theme -> CEWL_emmean_plot
CEWL_emmean_plot
```



This is **only** for lizards that were recaptured and measured more than once, so I think these changes may reflect true seasonality in their water balance.



### Export Multi-Panel Fig

Put in one plot together:

```{r}
ggarrange(CEWL_emmean_plot, 
          osml_emmean_plot, 
          mass_emmean_plot,
          SMI_emmean_plot,
          ncol = 4, nrow = 1,
#          ncol = 1, nrow = 3,
          labels = c("A", "B", "C", "D"),
          common.legend = TRUE,
          legend = "bottom"
          ) -> szn_change_multi_fig
#szn_change_multi_fig

ggsave(filename = "szn_change_multi_fig.pdf",
       plot = szn_change_multi_fig,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       width = 200, height = 80
       #width = 80, height = 210
       )
```













## FULL Time*Sex Marginal Means

I'm going to do literally the same exact thing, but now for **all of my data**, not just for the data on lizards with repeat measurements.

Check n's:

```{r}
dat %>%
  group_by(month, sex_M_F) %>%
  summarise(n = n())
```




### Mass

```{r}
full_time_sex_diffs_mass <- lmerTest::lmer(data = dat, mass_g ~ month*sex_M_F +
                                        (1|individual_ID))
full_time_sex_diffs_mass_aov <- data.frame(anova(full_time_sex_diffs_mass, 
                                                 type = "1", ddf = "Kenward-Roger")) %>%
  mutate(explanatory = factor(row.names(.)))

# mean values
full_emmeans_mass <- data.frame(emmeans(full_time_sex_diffs_mass, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "Mass")

# stat diffs by month
full_time_emmeans_mass <- data.frame(emmeans(full_time_sex_diffs_mass, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "Mass")

# stat diffs by sex
full_sex_emmeans_mass <- data.frame(emmeans(full_time_sex_diffs_mass, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "Mass")
```




### Body Condition

```{r}
full_time_sex_diffs_SMI <- lmerTest::lmer(data = dat, SMI ~ month*sex_M_F +
                                        (1|individual_ID))
full_time_sex_diffs_SMI_aov <- data.frame(anova(full_time_sex_diffs_SMI, 
                                          type = "1", ddf = "Kenward-Roger")) %>%
  mutate(explanatory = factor(row.names(.)))

# mean values
full_emmeans_SMI <- data.frame(emmeans(full_time_sex_diffs_SMI, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "SMI")

# stat diffs by month
full_time_emmeans_SMI <- data.frame(emmeans(full_time_sex_diffs_SMI, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "SMI")

# stat diffs by sex
full_sex_emmeans_SMI <- data.frame(emmeans(full_time_sex_diffs_SMI, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "SMI")
```



### Osmolality


```{r}
full_time_sex_diffs_osml <- lmerTest::lmer(data = dat, 
                                      osmolality_mmol_kg_mean ~ month*sex_M_F +
                                        (1|individual_ID))
full_time_sex_diffs_osml_aov <- data.frame(anova(full_time_sex_diffs_osml, 
                                           type = "1", ddf = "Kenward-Roger")) %>%
  mutate(explanatory = factor(row.names(.)))

# mean values
full_emmeans_osml <- data.frame(emmeans(full_time_sex_diffs_osml, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "Osmolality")

# stat diffs by month
full_time_emmeans_osml <- data.frame(emmeans(full_time_sex_diffs_osml, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "Osmolality")

# stat diffs by sex
full_sex_emmeans_osml <- data.frame(emmeans(full_time_sex_diffs_osml, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "Osmolality")
```




### CEWL

```{r}
full_time_sex_diffs_CEWL <- lmerTest::lmer(data = dat, 
                                      CEWL_g_m2h ~ month*sex_M_F +
                                        (1|individual_ID))
full_time_sex_diffs_CEWL_aov <- data.frame(anova(full_time_sex_diffs_CEWL, 
                                           type = "1", ddf = "Kenward-Roger")) %>%
  mutate(explanatory = factor(row.names(.)))

# check covariates
full_time_sex_diffs_CEWL_check_covaries <- lmerTest::lmer(data = dat, 
                                                     CEWL_g_m2h ~ month*sex_M_F
                                         + msmt_temp_C + msmt_VPD_kPa +
                                        (1|individual_ID))
anova(full_time_sex_diffs_CEWL_check_covaries, 
      type = "1", ddf = "Kenward-Roger") # not important at all

# mean values
full_emmeans_CEWL <- data.frame(emmeans(full_time_sex_diffs_CEWL, 
                                   pairwise ~ month | sex_M_F)$emmean) %>%
  mutate(measurement = "CEWL")

# stat diffs by month
full_time_emmeans_CEWL <- data.frame(emmeans(full_time_sex_diffs_CEWL, 
                                   pairwise ~ month | sex_M_F)$contrasts) %>%
  mutate(measurement = "CEWL")

# stat diffs by sex
full_sex_emmeans_CEWL <- data.frame(emmeans(full_time_sex_diffs_CEWL, 
                                   pairwise ~ sex_M_F | month)$contrasts) %>%
  mutate(measurement = "CEWL")
```



### Check Assumptions

```{r}
plot(full_time_sex_diffs_CEWL)
plot(full_time_sex_diffs_osml)
plot(full_time_sex_diffs_mass)
```



### Merge dfs

```{r}
full_emmeans_all <- full_emmeans_CEWL %>%
  rbind(full_emmeans_osml) %>%
  rbind(full_emmeans_mass) %>%
  rbind(full_emmeans_SMI) %>%
  mutate(month_sex = factor(paste(month, sex_M_F),
                            levels = c("April Female", "April Male",
                                       "May Female", "May Male",
                                       "July Female", "July Male")))

# all pairwise stats together
full_emmean_time_stats <- full_time_emmeans_CEWL %>%
  rbind(full_time_emmeans_osml) %>%
  rbind(full_time_emmeans_mass) %>%
  rbind(full_time_emmeans_SMI)
full_emmean_sex_stats <- full_sex_emmeans_CEWL %>%
  rbind(full_sex_emmeans_osml) %>%
  rbind(full_sex_emmeans_mass) %>%
  rbind(full_sex_emmeans_SMI)

# anova stats
full_time_sex_diffs_all <- (full_time_sex_diffs_CEWL_aov) %>%
  rbind((full_time_sex_diffs_osml_aov)) %>%
  rbind((full_time_sex_diffs_mass_aov)) %>%
  rbind((full_time_sex_diffs_SMI_aov)) %>%
  mutate(response = factor(c(rep("CEWL", 3), rep("Plasma Osmolality", 3),
                      rep("Mass", 3), rep("Body Condition", 3))))
summary(full_time_sex_diffs_all)
```



### Export

```{r}
full_time_sex_diffs_4exp <- full_time_sex_diffs_all %>%
  mutate(seq_sum_of_squares = round(Sum.Sq, 0), 
         F_statistic = round(F.value, 2),
         DenDF = round(DenDF, 0),
         F_full = paste(F_statistic, " (", NumDF, ",", DenDF, ")", sep = ""),
         p_value = round(Pr..F., 4)) %>%
  dplyr::select(response, explanatory,
                seq_sum_of_squares,
                #NumDF, DenDF,
                #F_statistic, 
                F_full,
                p_value)
write.csv(full_time_sex_diffs_4exp, 
          "./results_statistics/hydric_Fstats_by_monthsex.csv")
```
























## FULL Plot MMeans Time*Sex

### Mass

```{r}
ggplot() +
  geom_jitter(data = dat, # dat or dat_repeats ?
             aes(x = month_sex,
                 y = mass_g,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
              position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = full_emmeans_all[full_emmeans_all$measurement == "Mass",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = full_emmeans_all[full_emmeans_all$measurement == "Mass",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                 color = sex_M_F),
                width = .4) +
  xlab("") + 
  ylab("Lizard Mass (g)") +
  # sex differences
  annotate(geom = "text", x = 5.5, y = 54, label = "*", size = 6) +
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 61, label = "AB", size = 3) +
  annotate(geom = "text", x = 4, y = 61, label = "A", size = 3) +
  annotate(geom = "text", x = 6, y = 61, label = "B", size = 3) +
  # time differences for FEMALES
  annotate(geom = "text", x = 1, y = 61, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 61, label = "b", size = 3) +
  annotate(geom = "text", x = 5, y = 61, label = "a", size = 3) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels_full) +
  ylim(19, 61) +
  savs_ggplot_theme -> full_mass_emmean_plot
full_mass_emmean_plot
```



### Body Condition

```{r}
ggplot() +
  geom_jitter(data = dat, # dat or dat_repeats ?
             aes(x = month_sex,
                 y = SMI,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
              position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = full_emmeans_all[full_emmeans_all$measurement == "SMI",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = full_emmeans_all[full_emmeans_all$measurement == "SMI",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                 color = sex_M_F),
                width = .4) +
  xlab("") + 
  ylab("Body Condition (g')") +
  # sex differences
  annotate(geom = "text", x = 1.5, y = 44, label = "*", size = 6) +
  annotate(geom = "text", x = 3.5, y = 44, label = "*", size = 6) +
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 49, label = "A", size = 3) +
  annotate(geom = "text", x = 4, y = 49, label = "A", size = 3) +
  annotate(geom = "text", x = 6, y = 49, label = "B", size = 3) +
  # time differences for FEMALES
  annotate(geom = "text", x = 1, y = 49, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 49, label = "b", size = 3) +
  annotate(geom = "text", x = 5, y = 49, label = "a", size = 3) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels_full) +
  ylim(19, 50) +
  savs_ggplot_theme -> full_SMI_emmean_plot
full_SMI_emmean_plot
```



### Osmolality

```{r}
ggplot() +
  geom_jitter(data = dat,  # dat or dat_repeats ?
             aes(x = month_sex,
                 y = osmolality_mmol_kg_mean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
              position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = full_emmeans_all[full_emmeans_all$measurement == "Osmolality",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = full_emmeans_all[full_emmeans_all$measurement == "Osmolality",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                 color = sex_M_F),
                width = .4) +
  xlab("") + 
  ylab(bquote('Osmolality (mmol '*kg^-1*')')) +
  # sex differences
  annotate(geom = "text", x = 1.5, y = 422, label = "*", size = 6) +
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 445, label = "AB", size = 3) +
  annotate(geom = "text", x = 4, y = 445, label = "A", size = 3) +
  annotate(geom = "text", x = 6, y = 445, label = "B", size = 3) +
  # time differences for FEMALES
  annotate(geom = "text", x = 1, y = 445, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 445, label = "a", size = 3) +
  annotate(geom = "text", x = 5, y = 445, label = "a", size = 3) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels_full) +
  scale_y_continuous(limits = c(320, 450), breaks = c(320, 360, 400, 440)) +
  savs_ggplot_theme -> full_osml_emmean_plot
full_osml_emmean_plot
```




### CEWL

```{r}
ggplot() +
  geom_jitter(data = dat, # dat or dat_repeats ?
             aes(x = month_sex,
                 y = CEWL_g_m2h,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 0.3,
             size = 0.9,
             position = position_jitter(height = 0, width = 0.2)) +
  geom_point(data = full_emmeans_all[full_emmeans_all$measurement == "CEWL",],
             aes(x = month_sex,
                 y = emmean,
                 shape = sex_M_F,
                 color = sex_M_F,
                 fill = sex_M_F),
             alpha = 1,
             size = 2) +
  geom_errorbar(data = full_emmeans_all[full_emmeans_all$measurement == "CEWL",],
                aes(x = month_sex,
                    y = emmean, 
                    ymin = lower.CL,
                    ymax = upper.CL,
                    color = sex_M_F),
                width = .4) +
  # sex differences
  annotate(geom = "text", x = 3.5, y = 20, label = "*", size = 6) +
  # time differences for MALES
  annotate(geom = "text", x = 2, y = 24, label = "A", size = 3) +
  annotate(geom = "text", x = 4, y = 24, label = "B", size = 3) +
  annotate(geom = "text", x = 6, y = 24, label = "A", size = 3) +
  # time differences for FEMALES = none
  annotate(geom = "text", x = 1, y = 24, label = "a", size = 3) +
  annotate(geom = "text", x = 3, y = 24, label = "a", size = 3) +
  annotate(geom = "text", x = 5, y = 24, label = "a", size = 3) +
  xlab("") + 
  ylab(bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  ylim(0, 25) +
  scale_color_manual(values = my_colors, name = "Sex") +
  scale_fill_manual(values = my_colors, name = "Sex") +
  scale_shape_manual(values = my_shapes_open, name = "Sex") +
  scale_x_discrete(labels = my_labels_full) +
  savs_ggplot_theme -> full_CEWL_emmean_plot
full_CEWL_emmean_plot
```





### Export Multi-Panel Fig

Put in one plot together:

```{r}
ggarrange(full_CEWL_emmean_plot, 
          full_osml_emmean_plot, 
          full_mass_emmean_plot,
          full_SMI_emmean_plot,
          #ncol = 4, nrow = 1,
          ncol = 1, nrow = 4,
          labels = c("A", "B", "C"),
          common.legend = TRUE,
          legend = "bottom"
          ) -> full_szn_change_multi_fig
#full_szn_change_multi_fig

ggsave(filename = "szn_change_multi_fig_full.pdf",
       plot = full_szn_change_multi_fig,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       #width = 200, height = 80
       width = 80, height = 210
       )
```




### Export Full & Repeats Multi-Panel Fig


```{r}
ggarrange(full_CEWL_emmean_plot, 
          full_osml_emmean_plot, 
          full_mass_emmean_plot,
          full_SMI_emmean_plot,
          CEWL_emmean_plot, 
          osml_emmean_plot, 
          mass_emmean_plot,
          SMI_emmean_plot,
          ncol = 4, nrow = 2,
#          ncol = 1, nrow = 3,
          labels = c("full", "full", "full", "full",
                     "rep", "rep", "rep", "rep"),
          common.legend = TRUE,
          legend = "bottom"
          ) -> full_vs_rep_szn_change_multi_fig
#full_vs_rep_szn_change_multi_fig

ggsave(filename = "szn_change_multi_fig_full_vs_rep.pdf",
       plot = full_vs_rep_szn_change_multi_fig,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       width = 210, height = 160
       #width = 80, height = 210
       )
```










# Reproductive Effort

Is reproductive effort related to water balance?


## Repeat Individuals

I have 9 females measured at both April and May time points. Let's see if their physiology changes consistently based on becoming gravid. 


First, calculate the change in CEWL, osmolality, and mass for each of those females between April to May.

```{r}
summary(dat_repeat_females)
fem_starting_values <- dat_repeat_females %>%
  # get the values for April measurements only 
  dplyr::filter(month == "April") %>%
  dplyr::select(individual_ID,
                CEWL_April = CEWL_g_m2h,
                osml_April = osmolality_mmol_kg_mean,
                mass_April = mass_g)
fem_change <- dat_repeat_females %>%
  # add starting values
  left_join(fem_starting_values, by = 'individual_ID') %>%
  # make sure we have complete data
  dplyr::filter(complete.cases(mass_April, CEWL_April)) %>%
  # calculate difference for each msmt versus in April
  mutate(mass_change_from_April = mass_g - mass_April,
         CEWL_change_from_April = CEWL_g_m2h - CEWL_April,
         osml_change_from_April = osmolality_mmol_kg_mean - osml_April)
summary(fem_change)
```



## Plot Indiv Change

```{r}
ggplot(fem_change) +
  aes(x = month,
      y = mass_change_from_April,
      group = individual_ID,
      color = gravid_Y_N
      ) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplot(fem_change) +
  aes(x = month,
      y = CEWL_change_from_April,
      group = individual_ID,
      color = gravid_Y_N
      ) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplot(fem_change) +
  aes(x = month,
      y = osml_change_from_April,
      group = individual_ID,
      color = gravid_Y_N
      ) +
  geom_point() +
  geom_line() +
  theme_bw()
boxplot(fem_change$osml_change_from_April)
boxplot(fem_change$mass_change_from_April)
```


On average, mass and osmolality both increase. There's only one lizard who was not gravid at both time points, so we can't compare change based on becoming gravid or not. I think it is reasonable to conclude that on average, gravid females gained mass and became less hydrated. But, we cannot say that this was DUE to gravidity, because we have nothing to compare it to


## Stats Indiv Change

Run stats tests on the change rates for only the females who were gravid in May, and use only the actual delta values, in their rows for May data.

```{r}
# check osml outlier range
fem_change %>%
  group_by(month) %>%
  summarise(mean(osml_change_from_April), sd(osml_change_from_April))

# for stats
fem_change_stats <- fem_change %>%
  # don't use the one female still non-gravid in May
  dplyr::filter(individual_ID != "F-12") %>%
  # only use the >resulting< change values
  dplyr::filter(month == "May") %>%
  # remove the crazy low outlier based on sd calculated above
  dplyr::filter(osml_change_from_April > (12-(29*2)))
```


ttests:

```{r}
t.test(fem_change_stats$CEWL_change_from_April)
t.test(fem_change_stats$osml_change_from_April)
t.test(fem_change_stats$mass_change_from_April)
```

Only mass change was significantly different from zero... But, if we got rid of the one super low osmolality value, the change in osmolality would also be significant. 



## Stats Likelihood

Does anything correlate with the likelihood of being/becoming gravid?


Start with trying each possible logistic model to predict gravidity:

```{r}
repro_logit1 <- glm(gravid_Y_N ~ 
                     mass_g, 
                   data = dat_repro_females, 
                   family = "binomial")
repro_logit2 <- glm(gravid_Y_N ~ 
                     SMI, 
                   data = dat_repro_females, 
                   family = binomial)
repro_logit3 <- glm(gravid_Y_N ~ 
                     osmolality_mmol_kg_mean, 
                   data = dat_repro_females, 
                   family = binomial)
repro_logit4 <- glm(gravid_Y_N ~ 
                     CEWL_g_m2h, 
                   data = dat_repro_females, 
                   family = binomial)
summary(repro_logit1)
summary(repro_logit2)
summary(repro_logit3)
summary(repro_logit4)
```




## Plot Likelihood Gravid ~ Mass

```{r}
dat_repro_females %>%
  # dummy-fy gravid variable
  mutate(gravid_Y_N = factor(gravid_Y_N, levels = c("Not Gravid", "Gravid"),
                             labels = c(0,1)),
         # goes 1-2, so fix to be 0-1
         gravid_Y_N = as.numeric(gravid_Y_N)-1) %>%
  # nooooowwwww I can plot and the stat_smooth will actually work
  ggplot() +
  aes(x = mass_g, 
      y = gravid_Y_N) +
  geom_jitter(alpha=.5, 
              position = position_jitter(height = 0.1, width = 0)
              ) +
  # must make response variable 0-1 in order for this to work
  stat_smooth(geom = "smooth",
              method = "glm", 
              formula = y~x,
              se = FALSE, 
              method.args = list(family = "quasibinomial")
              ) +
  ylab("") +
  scale_y_continuous(labels = c("Not Gravid", "", "", "", "Gravid"), 
                     limits = c(0,1)) +
  xlab("Body Mass (g)") +
  savs_ggplot_theme
```





# July Only

Was hydric physiology different between lizards that were estivating vs above ground still when we removed their collars in July?

## Get Estivation Onset 

I get this from the 'analysis_2' data/code:

```{r}
Serials <- c("229-069", "229-078", "229-079", "245-742", "245-745", "245-748",
             "245-752", "245-753", "245-761", "252-881", "252-883")
estivating_yet <- c("Yes", "Yes", "Maybe", "Yes", "Maybe", 
                    "Yes", "Yes", "Yes", "Maybe", "Yes", "Yes")
est_df <- data.frame(radio_collar_serial = Serials,
                     est_status = estivating_yet)
```

Now add that info to the hydric msmts:


```{r}
dat_July <- dat %>%
  dplyr::filter(capture_date == "2021-07-14") %>%
  left_join(est_df, by = 'radio_collar_serial') %>%
  dplyr::filter(complete.cases(est_status))
```


## Stats of Diffs

```{r}
summary(lm(data = dat_July, CEWL_g_m2h ~ est_status))
summary(lm(data = dat_July, osmolality_mmol_kg_mean ~ est_status))
hist(dat_July$osmolality_mmol_kg_mean)
hist(dat_July$CEWL_g_m2h)
```

There is absolutely NO difference between lizards that were estivating or not when we measured them in July. And, which ones were estivating is just an estimate. I wouldn't trust it because we definitely lasso-ed a few lizards, but the microhabitat observations don't have any above-ground notes for any of the lizards. 


















































