---
title: "Species Comparison of Water Balance"
author: "Savannah Weaver"
output:
  rmarkdown::html_vignette:
      highlight: tango
      thumbnails: FALSE
      toc: TRUE
      toc_depth: 3
---


# Packages

```{r setup, echo=T, results='hide', message=FALSE}
if (!require("tidyverse")) install.packages("tidyverse") 
library("tidyverse") # for tidy/dplyr workflow
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2") # plots!
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # for multi-ggplot figs
if (!require("emmeans")) install.packages("emmeans")
library("emmeans") # marginal means & confints
if (!require("RColorBrewer")) install.packages("RColorBrewer")
library("RColorBrewer") # pretty colors
```









# Set Colors & Themes

```{r}
savs_ggplot_theme <- theme_classic() + 
                     theme(text = element_text(color = "black", 
                                               family = "sans", 
                                               size = 12),
                           axis.text = element_text(color = "black", 
                                                    family = "sans", 
                                                    size = 8),
                           legend.text = element_text(color = "black", 
                                                      family = "sans", 
                                                      size = 8),
                           legend.text.align = 0,
                           legend.position = "bottom"
                           )
  
# custom colors & shapes for WFL vs BNLL
Scelop <- brewer.pal(11, "RdYlBu")[c(9)]
Gambel <- brewer.pal(11, "RdYlBu")[c(2)]
my_colors <- c(Gambel, Scelop)
my_shapes <- c(21, 23)

# custom, consistent legend labels
# extra spaces make room for species photos in legend 
my_labels <- c(expression(italic("Gambelia sila                                                              ")),
               expression(italic("Sceloporus occidentalis                                      ")))
```










# Background

Because there are so few data measuring CEWL with the same methods we use, we decided to compare BNLL to two Mediterranean lizards that we've measured using the exact same methods.

look at CEWL-osml distribution for BNLL & WFL














# Load Data




## *Gambelia sila* Data

```{r load Gambelia data}
Gambelia <- read_rds("./data/G_sila_clean_full_dat.RDS") %>% 
  dplyr::select(date = capture_date, individual_ID, 
                CEWL_g_m2h,
                msmt_temp_C,
                msmt_RH_percent,
                osmolality_mmol_kg = osmolality_mmol_kg_mean,
                sex_M_F, mass_g, SVL_mm) %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg, msmt_temp_C)) %>%
  # add column for species
  mutate(species = "Gambelia sila")
summary(Gambelia)
```






## *Sceloporus occidentalis* Data

This is from the experiment I ran summer 2021 (submitted to JEB for publication).

```{r load Sceloporus data}
Sceloporus <- read_rds("./data/Sceloporus_2021_exp.RDS") %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>% 
  dplyr::select(date = capture_date, individual_ID, 
                CEWL_g_m2h = CEWL_g_m2h_mean,
                msmt_temp_C,
                msmt_RH_percent,
                osmolality_mmol_kg = osmolality_mmol_kg_mean,
                mass_g, SVL_mm) %>%
  # format variables
  mutate(sex_M_F = factor("Male")) %>% 
  # add column for species
  mutate(species = "Sceloporus occidentalis")
summary(Sceloporus)
names(Sceloporus)
```





## Merge Dataframes

### S.o. and G.s.

```{r merge}
dat_CA_liz <- Gambelia %>%
  rbind(Sceloporus) %>%
  mutate(species = factor(species),
         # calculate VPD based on Campbell & Norman 1998
         e_s_kPa = 0.611 * exp((17.502*msmt_temp_C)/(msmt_temp_C + 240.97)),
         msmt_VPD_kPa = e_s_kPa*(1 - (msmt_RH_percent/100))
         ) 
summary(dat_CA_liz)
```









# Check Distributions


## Covariates

First, I need to check that the conditions measurements were taken in mostly overlapped for the two species.

The temperature and VPD figures will actually be used in publication. I'll make nicer versions later.

```{r}
ggplot(dat_CA_liz) +
  aes(x = msmt_temp_C,
      y = CEWL_g_m2h,
      color = species) +
  geom_point()
ggplot(dat_CA_liz) +
  aes(x = msmt_RH_percent,
      y = CEWL_g_m2h,
      color = species) +
  geom_point()
ggplot(dat_CA_liz) +
  aes(x = msmt_VPD_kPa,
      y = CEWL_g_m2h,
      color = species) +
  geom_point()
```



## Body Sizes

Also get stats on how different the two species are.

```{r}
ggplot(dat_CA_liz) +
  aes(x = SVL_mm,
      y = mass_g,
      color = species) +
  geom_point()
dat_CA_liz %>%
  group_by(species, sex_M_F) %>% 
  summarise(mean_SVL_mm = mean(SVL_mm),
            mean_mass = mean(mass_g))
```










# Marginal Mean Differences

## osmolality

```{r}
sp_lm_osml <- lm(data = dat_CA_liz, osmolality_mmol_kg ~ species)
osml_emmean <- data.frame(emmeans(sp_lm_osml, "species")) %>%
  rename(osml_emmean = emmean,
         osml_lower = lower.CL,
         osml_upper = upper.CL)
pairs(emmeans(sp_lm_osml, "species"))
```


## CEWL

```{r}
sp_lm_CEWL <- lm(data = dat_CA_liz, CEWL_g_m2h ~ species)
CEWL_emmean <- data.frame(emmeans(sp_lm_CEWL, "species")) %>%
  rename(CEWL_emmean = emmean,
         CEWL_lower = lower.CL,
         CEWL_upper = upper.CL)
pairs(emmeans(sp_lm_CEWL, "species"))
```

put marginal means together

```{r}
both_emmeans <- osml_emmean %>%
  left_join(CEWL_emmean, by = 'species')
both_emmeans
```



## temp

```{r}
sp_lm_msmt_temp <- lm(data = dat_CA_liz, msmt_temp_C ~ species)
temp_emmean <- data.frame(emmeans(sp_lm_msmt_temp, "species")) %>%
  rename(temp_emmean = emmean,
         temp_lower = lower.CL,
         temp_upper = upper.CL)
pairs(emmeans(sp_lm_msmt_temp, "species"))

CEWL_temp_emmeans <- temp_emmean %>%
  left_join(CEWL_emmean, by = 'species')
CEWL_temp_emmeans
```


## VPD

```{r}
sp_lm_msmt_VPD <- lm(data = dat_CA_liz, msmt_VPD_kPa ~ species)
VPD_emmean <- data.frame(emmeans(sp_lm_msmt_VPD, "species")) %>%
  rename(VPD_emmean = emmean,
         VPD_lower = lower.CL,
         VPD_upper = upper.CL)
pairs(emmeans(sp_lm_msmt_VPD, "species"))

CEWL_VPD_emmeans <- VPD_emmean %>%
  left_join(CEWL_emmean, by = 'species')
CEWL_VPD_emmeans
```







# Figures




## Temp Covariate

CEWL ~ temperature:

```{r}
ggplot() +
  geom_point(data = dat_CA_liz,
             aes(x = msmt_temp_C,
                  y = CEWL_g_m2h,
                  color = species,
                  shape = species,
                  fill = species),
      alpha = 0.3,
      size = 0.9) +
  geom_point(data = CEWL_temp_emmeans,
             aes(x = temp_emmean,
                 y = CEWL_emmean,
                 shape = species,
                 fill = species),
             size = 3,
             alpha = 1,
             color = "black") +
  scale_color_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_fill_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_shape_manual(values = my_shapes, name = NULL, labels = my_labels) +
  #scale_x_continuous(limits = c(1.37,4.32)) +
  scale_y_continuous(limits = c(0,35)) +
  labs(x = "Temperature (°C)",
       y = bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  savs_ggplot_theme + 
  theme(
    legend.margin = margin(25,0,25,0, unit = "pt")
  )-> CEWL_temp_check
CEWL_temp_check
```




## VPD Covariate

```{r}
ggplot() +
  geom_point(data = dat_CA_liz,
             aes(x = msmt_VPD_kPa,
                  y = CEWL_g_m2h,
                  color = species,
                  shape = species,
                  fill = species),
      alpha = 0.3,
      size = 0.9) +
  geom_point(data = CEWL_VPD_emmeans,
             aes(x = VPD_emmean,
                 y = CEWL_emmean,
                 shape = species,
                 fill = species),
             size = 3,
             alpha = 1,
             color = "black") +
  scale_color_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_fill_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_shape_manual(values = my_shapes, name = NULL, labels = my_labels) +
  scale_x_continuous(limits = c(1.37,4.32)) +
  scale_y_continuous(limits = c(0,35)) +
  labs(x = "VPD (kPa)",
       y = bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  savs_ggplot_theme + 
  theme(
    legend.margin = margin(25,0,25,0, unit = "pt")
  )-> CEWL_VPD_check
CEWL_VPD_check
```








## CEWL ~ Osml



```{r}
ggplot() +
  # all data
  geom_point(data = dat_CA_liz,
             aes(x = osmolality_mmol_kg,
                 y = CEWL_g_m2h,
                 fill = species,
                 color = species,
                 shape = species),
             alpha = 0.3) + 
  # mean points
  geom_point(data = both_emmeans,
             aes(x = osml_emmean,
                 y = CEWL_emmean,
                 fill = species,
                 shape = species),
             size = 5,
             color = "black") + 
  
  scale_fill_manual(values = my_colors, labels = my_labels, name = NULL) +
  scale_color_manual(values = my_colors, labels = my_labels, name = NULL) +
  scale_shape_manual(values = my_shapes, labels = my_labels, name = NULL) +
  scale_x_continuous(limits = c(300, 450)) +
  scale_y_continuous(limits = c(0, 35)) +
  labs(x = bquote('Plasma Osmolality (mmol '*kg^-1*')'),
       y = bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  savs_ggplot_theme + 
  theme(
    legend.margin = margin(25,0,25,0, unit = "pt")
  )-> Scelop_BNLL_pts
Scelop_BNLL_pts
```




## Export Grouped Figure


```{r}
Scelop_BNLL_pts_pub <- Scelop_BNLL_pts +
  theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))
CEWL_temp_check_pub <- CEWL_temp_check +
  theme(legend.position = "none",
        plot.margin = margin(t = 0, r = 0, b = 6, l = 3, unit = "pt")) +
  labs(y = NULL)
CEWL_VPD_check_pub <- CEWL_VPD_check +
  theme(legend.position = "none",
        plot.margin = margin(t = 0, r = 0, b = 6, l = 3, unit = "pt")) +
  labs(y = NULL)

ggarrange(Scelop_BNLL_pts_pub, 
          ggarrange(CEWL_temp_check_pub, CEWL_VPD_check_pub,
                    ncol = 1, nrow = 2, 
                    labels = c("(b)", "(c)"),
                    font.label = list(color = "black", face = "bold", size = 10),
                    label.y = 1.03,
                    label.x = -0.05),
          ncol = 2, nrow = 1,
          widths = c(2, 1),
          labels = c("(a)"),
          label.y = 1.015, label.x = 0.03,
          font.label = list(color = "black", face = "bold", size = 10),
          common.legend = TRUE,
          legend = "bottom"
          ) -> CEWL_BNLL_WFL_comparison

ggsave(filename = "Scelop_BNLL_compare.pdf",
       plot = CEWL_BNLL_WFL_comparison,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       width = 160, height = 100)
```
















