---
title: "Species Comparison of Water Balance"
author: "Savannah Weaver"
output: 
  rmdformats::html_clean:
    highlight: tango
    thumbnails: FALSE
    toc: TRUE
    toc_depth: 3
---


# Packages

```{r setup, echo=T, results='hide', message=FALSE}
if (!require("tidyverse")) install.packages("tidyverse") 
library("tidyverse")# for dplyr workflow of calculations
if (!require("UsingR")) install.packages("UsingR") 
library("UsingR")
if (!require("lme4")) install.packages("lme4") 
library("lme4")
if (!require("lmerTest")) install.packages("lmerTest") 
library("lmerTest")
if (!require("broom")) install.packages("broom")
library("broom") # lmer model export
if (!require("broom.mixed")) install.packages("broom.mixed")
library("broom.mixed") # lmer model export
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # for multi-ggplot figs
if (!require("Hmisc")) install.packages("Hmisc")
library("Hmisc") # correlation matrix
if (!require("emmeans")) install.packages("emmeans")
library("emmeans") # marginal means & confints
if (!require("RColorBrewer")) install.packages("RColorBrewer")
library("RColorBrewer") # marginal means & confints
```









# Set Colors & Themes

```{r}
savs_ggplot_theme <- theme_classic() + 
                     theme(text = element_text(color = "black", 
                                               family = "sans", 
                                               size = 12),
                           axis.text = element_text(color = "black", 
                                                    family = "sans", 
                                                    size = 8),
                           legend.text = element_text(color = "black", 
                                                      family = "sans", 
                                                      size = 8),
                           legend.text.align = 0,
                           legend.position = "bottom"
                           )
  
# custom colors & shapes for WFL vs BNLL
Scelop <- brewer.pal(11, "RdYlBu")[c(9)]
Gambel <- brewer.pal(11, "RdYlBu")[c(2)]
my_colors <- c(Gambel, Scelop)
my_shapes <- c(21, 23)

# custom, consistent legend labels
my_labels <- c(expression(italic("Gambelia sila")),
               expression(italic("Sceloporus occidentalis")))
```










# Background

Because there are so few data measuring CEWL with the same methods we use, we decided to compare BNLL to two Mediterranean lizards that we've measured using the exact same methods.

look at CEWL-osml distribution for BNLL & WFL
look at CEWL diffs for BNLL, WFL, and AWL















# Load Data




## *Gambelia sila* Data

```{r load Gambelia data}
BNLL_dat <- read_rds("./data/G_sila_clean_full_dat.RDS") %>% 
  dplyr::select(date = capture_date, individual_ID, 
                CEWL_g_m2h,
                msmt_temp_C,
                msmt_RH_percent,
                osmolality_mmol_kg = osmolality_mmol_kg_mean,
                sex_M_F, mass_g, SVL_mm) %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg, msmt_temp_C)) %>%
  # add column for species
  mutate(species = "Gambelia sila",
         group = "G")
summary(BNLL_dat)
```






## *Sceloporus occidentalis* Data


```{r load Sceloporus data}
Sceloporus_study1 <- read_rds("./data/Sceloporus_2021_CURE.RDS") %>%
  dplyr::filter(region == "Dorsum") %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>% 
  dplyr::select(date, individual_ID, 
                CEWL_g_m2h = TEWL_g_m2h,
                msmt_temp_C = ambient_temp_C,
                msmt_RH_percent = ambient_RH_percent,
                osmolality_mmol_kg,
                sex_M_F, mass_g, SVL_mm) %>%
  mutate(sex_M_F = factor(sex_M_F, labels = c("Female", "Male")),
         group = "A")
Sceloporus_study2 <- read_rds("./data/Sceloporus_2021_exp.RDS") %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg_mean)) %>% 
  dplyr::select(date = capture_date, individual_ID, 
                CEWL_g_m2h = CEWL_g_m2h_mean,
                msmt_temp_C,
                msmt_RH_percent,
                osmolality_mmol_kg = osmolality_mmol_kg_mean,
                mass_g, SVL_mm) %>%
  mutate(sex_M_F = factor("Male"),
         group = "B")
  
Sceloporus <- Sceloporus_study2 %>% 
  #rbind(Sceloporus_study1) %>%
  # rbind 2022 data?
  # add column for species
  mutate(species = "Sceloporus occidentalis")
summary(Sceloporus)
```



## *Podarcis* Data

```{r load Podarcis data}
Podarcis <- read_csv("./data/Podarcis_2022.csv") %>%
  dplyr::filter(Body == "Dorsum") %>%
  dplyr::select(date = Date, individual_ID = ID, 
                #sex_M_F = Sex, mass_g = Mass, SVL_mm = SVL,
                CEWL_g_m2h = CEWL_g_m2h_mean,
                msmt_temp_C, msmt_RH_percent) %>%
  mutate(date = as.Date(date, format = "%m/%d/%y"),
         individual_ID = factor(individual_ID),
         #sex_M_F = factor(sex_M_F),
         species = "Podarcis erhardii",
         # calculate VPD based on Campbell & Norman 1998
         e_s_kPa = 0.611 * exp((17.502*msmt_temp_C)/(msmt_temp_C + 240.97)),
         msmt_VPD_kPa = e_s_kPa*(1 - (msmt_RH_percent/100)))
summary(Podarcis)
```


## Merge S.o. and G.s.

```{r merge}
dat_CA_liz <- BNLL_dat %>%
  rbind(Sceloporus) %>%
  mutate(species = factor(species),
         # calculate VPD based on Campbell & Norman 1998
         e_s_kPa = 0.611 * exp((17.502*msmt_temp_C)/(msmt_temp_C + 240.97)),
         msmt_VPD_kPa = e_s_kPa*(1 - (msmt_RH_percent/100))
         ) 
summary(dat_CA_liz)
```











# Check Distributions


## Covariates

First, I need to check that the conditions measurements were taken in mostly overlapped for the two species.

The temperature and VPD figures will actually be used in publication. I'll make nicer versions later.

```{r}
ggplot(dat_CA_liz) +
  aes(x = msmt_temp_C,
      y = CEWL_g_m2h,
      color = species) +
  geom_point()
ggplot(dat_CA_liz) +
  aes(x = msmt_RH_percent,
      y = CEWL_g_m2h,
      color = species) +
  geom_point()
ggplot(dat_CA_liz) +
  aes(x = msmt_VPD_kPa,
      y = CEWL_g_m2h,
      color = species) +
  geom_point()
```



## Body Sizes

Also get stats on how different the two species are.

```{r}
ggplot(dat_CA_liz) +
  aes(x = SVL_mm,
      y = mass_g,
      color = group) +
  geom_point()
dat_CA_liz %>%
  group_by(species, sex_M_F) %>% 
  summarise(mean_SVL_mm = mean(SVL_mm),
            mean_mass = mean(mass_g))
```










# Marginal Mean Differences

## osmolality

```{r}
sp_lm_osml <- lm(data = dat_CA_liz, osmolality_mmol_kg ~ species)
osml_emmean <- data.frame(emmeans(sp_lm_osml, "species")) %>%
  rename(osml_emmean = emmean,
         osml_lower = lower.CL,
         osml_upper = upper.CL)
pairs(emmeans(sp_lm_osml, "species"))
```


## CEWL

```{r}
sp_lm_CEWL <- lm(data = dat_CA_liz, CEWL_g_m2h ~ species)
CEWL_emmean <- data.frame(emmeans(sp_lm_CEWL, "species")) %>%
  rename(CEWL_emmean = emmean,
         CEWL_lower = lower.CL,
         CEWL_upper = upper.CL)
pairs(emmeans(sp_lm_CEWL, "species"))
```

put marginal means together

```{r}
both_emmeans <- osml_emmean %>%
  left_join(CEWL_emmean, by = 'species')
both_emmeans
```



## temp

```{r}
sp_lm_msmt_temp <- lm(data = dat_CA_liz, msmt_temp_C ~ species)
temp_emmean <- data.frame(emmeans(sp_lm_msmt_temp, "species")) %>%
  rename(temp_emmean = emmean,
         temp_lower = lower.CL,
         temp_upper = upper.CL)
pairs(emmeans(sp_lm_msmt_temp, "species"))

CEWL_temp_emmeans <- temp_emmean %>%
  left_join(CEWL_emmean, by = 'species')
CEWL_temp_emmeans
```


## VPD

```{r}
sp_lm_msmt_VPD <- lm(data = dat_CA_liz, msmt_VPD_kPa ~ species)
VPD_emmean <- data.frame(emmeans(sp_lm_msmt_VPD, "species")) %>%
  rename(VPD_emmean = emmean,
         VPD_lower = lower.CL,
         VPD_upper = upper.CL)
pairs(emmeans(sp_lm_msmt_VPD, "species"))

CEWL_VPD_emmeans <- VPD_emmean %>%
  left_join(CEWL_emmean, by = 'species')
CEWL_VPD_emmeans
```







# Figures




## Temp Covariate

CEWL ~ temperature:

```{r}
ggplot() +
  geom_point(data = dat_CA_liz,
             aes(x = msmt_temp_C,
                  y = CEWL_g_m2h,
                  color = species,
                  shape = species,
                  fill = species),
      alpha = 0.3,
      size = 0.9) +
  geom_point(data = CEWL_temp_emmeans,
             aes(x = temp_emmean,
                 y = CEWL_emmean,
                 shape = species,
                 fill = species),
             size = 3,
             alpha = 1,
             color = "black") +
  scale_color_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_fill_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_shape_manual(values = my_shapes, name = NULL, labels = my_labels) +
  #scale_x_continuous(limits = c(1.37,4.32)) +
  scale_y_continuous(limits = c(0,35)) +
  labs(x = "Temperature (°C)",
       y = bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  savs_ggplot_theme -> CEWL_temp_check
```




## VPD Covariate

```{r}
ggplot() +
  geom_point(data = dat_CA_liz,
             aes(x = msmt_VPD_kPa,
                  y = CEWL_g_m2h,
                  color = species,
                  shape = species,
                  fill = species),
      alpha = 0.3,
      size = 0.9) +
  geom_errorbar(data = CEWL_VPD_emmeans,
             aes(x = VPD_emmean,
                 ymin = CEWL_lower,
                 ymax = CEWL_upper,
                 color = species),
             width = 0.1) +
  geom_errorbarh(data = CEWL_VPD_emmeans,
             aes(y = CEWL_emmean,
                 xmin = VPD_lower,
                 xmax = VPD_upper,
                 color = species),
             height = 2
             ) +
  geom_point(data = CEWL_VPD_emmeans,
             aes(x = VPD_emmean,
                 y = CEWL_emmean,
                 #color = species,
                 shape = species,
                 fill = species),
             size = 3,
             alpha = 1,
             color = "black") +
  scale_color_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_fill_manual(values = my_colors, name = NULL, labels = my_labels) +
  scale_shape_manual(values = my_shapes, name = NULL, labels = my_labels) +
  scale_x_continuous(limits = c(1.37,4.32)) +
  scale_y_continuous(limits = c(0,35)) +
  labs(x = "VPD (kPa)",
       y = bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  savs_ggplot_theme -> CEWL_VPD_check
```








## CEWL ~ Osml



```{r}
ggplot() +
  # all data
  geom_point(data = dat_CA_liz,
             aes(x = osmolality_mmol_kg,
                 y = CEWL_g_m2h,
                 fill = species,
                 color = species,
                 shape = species),
             alpha = 0.3) + 
  # group
  #stat_ellipse(data = dat_CA_liz,
   #          aes(x = osmolality_mmol_kg,
    #              y = CEWL_g_m2h,
     #             color = species),
      #       size = 0.5,
       #        show.legend = FALSE) +
  geom_errorbar(data = both_emmeans,
                aes(x = osml_emmean,
                    ymin = CEWL_lower,
                    ymax = CEWL_upper,
                    color = species),
                width = 7) +
  geom_errorbarh(data = both_emmeans,
                aes(y = CEWL_emmean,
                    xmin = osml_lower,
                    xmax = osml_upper,
                    color = species),
                height = 2) +
  # mean points
  geom_point(data = both_emmeans,
             aes(x = osml_emmean,
                 y = CEWL_emmean,
                 fill = species,
                 shape = species),
             size = 5,
             color = "black") + 
  
  scale_fill_manual(values = my_colors, labels = my_labels, name = NULL) +
  scale_color_manual(values = my_colors, labels = my_labels, name = NULL) +
  scale_shape_manual(values = my_shapes, labels = my_labels, name = NULL) +
  scale_x_continuous(limits = c(300, 450)) +
  scale_y_continuous(limits = c(0, 35)) +
  labs(x = bquote('Plasma Osmolality (mmol '*kg^-1*')'),
       y = bquote('CEWL (g '*m^-2*' '*h^-1*')')) +
  savs_ggplot_theme -> Scelop_BNLL_pts
Scelop_BNLL_pts
```




## Export Grouped Figure


```{r}
Scelop_BNLL_pts_pub <- Scelop_BNLL_pts +
  theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))
CEWL_temp_check_pub <- CEWL_temp_check +
  theme(legend.position = "none",
        plot.margin = margin(t = 0, r = 0, b = 6, l = 3, unit = "pt")) +
  labs(y = NULL)
CEWL_VPD_check_pub <- CEWL_VPD_check +
  theme(legend.position = "none",
        plot.margin = margin(t = 0, r = 0, b = 6, l = 3, unit = "pt")) +
  labs(y = NULL)

ggarrange(Scelop_BNLL_pts_pub, 
          ggarrange(CEWL_temp_check_pub, CEWL_VPD_check_pub,
                    ncol = 1, nrow = 2, labels = c("B", "C"),
                    label.y = 1.03,
                    label.x = -0.05),
          ncol = 2, nrow = 1,
          widths = c(2, 1),
          labels = c("A"),
          label.y = 1.015, label.x = 0.03,
          common.legend = TRUE,
          legend = "bottom"
          ) -> CEWL_BNLL_WFL_comparison

ggsave(filename = "Scelop_BNLL_compare.pdf",
       plot = CEWL_BNLL_WFL_comparison,
       path = "./results_figures",
       device = "pdf",
       dpi = 600,
       units = "mm",
       width = 160, height = 100)
```








