---
title: "BNLL Physiology - General Data Wrangling"
author: "Savannah Weaver"
output: 
  rmarkdown::html_vignette:
    highlight: tango
    thumbnails: FALSE
    toc: TRUE
    toc_depth: 3
---


# Packages

```{r setup, include = TRUE}
`%nin%` = Negate(`%in%`)
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse") # workflow and plots
if (!require("rmdformats")) install.packages("rmdformats")
library("rmdformats") # clean html R markdown format
library(stringr)
```


# Background and Goals

Blood was drawn from the postorbital sinus of Blunt-nosed Leopard Lizards (*Gambelia sila*) between April - July 2021. After centrifuging and separating, plasma was run on a VAPRO vapor pressure osmometer in 1-3 replicates, when plasma volume allowed. In this R script, I check the distribution of replicates, omit outliers, and average remaining replicates. The final values will be more precise and accurate estimates of the true plasma osmolality for each lizard, and those values will be used in the analyses R script file. Please refer to **doi:** for the published scientific paper  and full details.


# Load Data



## CEWL

```{r load CEWL data}
CEWL <- read_rds("./data/CEWL_dat_all_clean.RDS") %>%
  # fix some upper/lowercase issues
  mutate(individual_ID = str_to_title(tolower(individual_ID)),
         # M-03a was originally W-029, so we need to reassign
         individual_ID = replace(individual_ID, individual_ID == "W-029", "M-03a"),
         individual_ID = factor(individual_ID))

summary(CEWL)
unique(CEWL$individual_ID)
```


## Osmolality

```{r load osml data}
osmolality <- read_rds("./data/osml_means_clean.RDS")
summary(osmolality)
```


## Ultrasounding on Eggs

```{r get ultrasound data}
eggs <- read.csv("./data/females_ultrasounding.csv") %>%
  mutate(ultrasound_date = as.Date(ultrasound_date, format = "%m/%d/%y"),
         individual_ID = factor(individual_ID),
         gravid_Y_N = factor(gravid_Y_N,
                             levels = c("Y", "N"),
                             labels = c("Gravid", "Not Gravid")),
         stage = str_trim(stage)) %>%
  left_join(read.csv("./data/female_egg_dvmt_key.csv"), 
            by = c('stage' = 'name')) %>%
  mutate(stage = factor(stage,
                        levels = c("small round", "large round", "soft",
                                   "soft oblong", "firm oblong", "hard oblong")),
         dev_point = stage.y) %>%
  dplyr::select(-X, -stage.y) %>%
  mutate(month = factor(substr(as.character(ultrasound_date), 7, 7),
                          labels = c("April", "May")))

summary(eggs)
```


## Radio-Collared Lizards


```{r get data collared lizards}
lizards_collared <- read.csv("./data/collared_measurements.csv") %>%
  mutate(capture_process_date_time = as.POSIXct(paste(capture_process_date, 
                                                      capture_time, 
                                                      sep = " "), 
                                                format = "%m/%d/%y %H:%M"),
                capture_process_date = as.Date(capture_process_date, 
                                                  format = "%m/%d/%y"),
                SVL_mm = as.numeric(SVL_mm),
                pre_tmt_mass_g = as.numeric(pre_tmt_mass_g),
                pre_mass_w_collar_Y_N = as.factor(pre_mass_w_collar_Y_N),
                post_tmt_mass_g = as.numeric(post_tmt_mass_g),
                post_mass_w_collar_Y_N = as.factor(post_mass_w_collar_Y_N),
                treatment_date_time = as.POSIXct(paste(treatment_date, 
                                                       processing_time, 
                                                       sep = " "), 
                                                 format = "%m/%d/%y %H:%M"),
                tmt_start_time = as.POSIXct(tmt_start_time, format = "%H:%M"),
                tmt_end_time = as.POSIXct(tmt_end_time, format = "%H:%M"),
                hematocrit_percent = as.numeric(hematocrit_percent),
                cloacal_temp_C = as.numeric(cloacal_temp_C)
                ) %>%
  # add tmt group info
  left_join(read.csv("./data/collared_identification.csv"),
            by = 'individual_ID') %>%
                # format data classes
  dplyr::mutate(individual_ID = as.factor(individual_ID),
                sex_M_F = as.factor(sex),
                tmt = as.factor(tmt),
                radio_collar_mass_g = as.numeric(radio_collar_mass_g),
                #statistics
                capture_to_msmt = as.numeric(treatment_date_time - capture_process_date_time),
                tmt_time_elapsed = as.numeric(tmt_end_time - tmt_start_time),
                actual_pre_tmt_mass_g = pre_tmt_mass_g - radio_collar_mass_g*(pre_mass_w_collar_Y_N == "Y"),
                actual_post_tmt_mass_g = post_tmt_mass_g - radio_collar_mass_g*(post_mass_w_collar_Y_N == "Y"),
                tmt_mass_change_g = actual_post_tmt_mass_g - actual_pre_tmt_mass_g,
                radio_collar_percent_mass_g = radio_collar_mass_g/actual_pre_tmt_mass_g
                ) %>%
  dplyr::select(capture_date_time = capture_process_date_time,
                capture_date = capture_process_date,
                treatment_date_time,
                radio_collar_serial, 
                PIT_tag_ID, 
                individual_ID,
                sex_M_F, 
                mass_g = actual_pre_tmt_mass_g,
                SVL_mm,
                tmt,
                tmt_time_elapsed,
                tmt_mass_change_g,
                radio_collar_mass_g,
                radio_collar_percent_mass_g,
                capture_to_msmt, 
                hematocrit_percent,
                cloacal_temp_C
                )

summary(lizards_collared)
unique(lizards_collared$radio_collar_serial)
```


## Widespread Lizards


```{r get data widespread lizards}
lizards_widespread <- read.csv("./data/widespread_measurements.csv") %>%
                # format data classes
  dplyr::mutate(capture_date_time = as.POSIXct(paste(date, 
                                                      capture_time, 
                                                      sep = " "), 
                                                format = "%m/%d/%y %H:%M"),
                capture_date = as.Date(date, format = "%m/%d/%y"),
                capture_time = as.POSIXct(capture_time, format = "%H:%M"),
                processing_time = as.POSIXct(processing_time, format = "%H:%M"),
                individual_ID = as.factor(individual_ID),
                SVL_mm = as.numeric(SVL_mm),
                mass_g = as.numeric(mass_g),
                sex_M_F = as.factor(sex_M_F),
                hematocrit_percent = as.numeric(hematocrit_percent),
                cloacal_temp_C = as.numeric(cloacal_temp_C),
                # statistics
                capture_to_msmt = as.numeric(processing_time - capture_time),
                # add things to match with treatment lizards
                tmt = as.factor("No Tmt"),
                radio_collar_serial = NA,
                tmt_mass_change_g = 0
                ) %>%
  dplyr::select(capture_date_time,
                capture_date,
                PIT_tag_ID, 
                individual_ID,
                sex_M_F, 
                mass_g,
                SVL_mm,
                capture_to_msmt, 
                hematocrit_percent,
                cloacal_temp_C,
                radio_collar_serial,
                tmt, tmt_mass_change_g
                )

summary(lizards_widespread)
```





# Methods Stats

How many lizards were captured/measured when?

```{r n lizards}
# lizards that were radio-collared
length(unique(lizards_collared$individual_ID))
# other lizards
length(unique(lizards_widespread$individual_ID))
# total lizards
length(unique(lizards_collared$individual_ID)) + length(unique(lizards_widespread$individual_ID))

# by date
lizards_widespread %>%
  group_by(capture_date) %>%
  summarise(n = n())
# by date
lizards_collared %>%
  group_by(capture_date) %>%
  summarise(n = n())

# which widespread were recaptured and when
lizards_widespread %>%
  group_by(individual_ID) %>%
  summarise(n = n()) %>%
  dplyr::filter(n > 1) -> recap_W
lizards_widespread %>% 
  dplyr::filter(individual_ID %in% recap_W$individual_ID) %>%
  arrange(individual_ID)
# which collared were recaptured and when
lizards_collared %>%
  group_by(individual_ID) %>%
  summarise(n = n()) %>%
  dplyr::filter(n > 1) -> recap_C
lizards_collared %>% 
  dplyr::filter(individual_ID %in% recap_C$individual_ID) %>%
  arrange(individual_ID)

# check things
lizards_collared %>% 
  dplyr::filter(individual_ID %nin% recap_C$individual_ID)
lizards_widespread %>% dplyr::filter(individual_ID == "W-029")
lizards_collared %>% dplyr::filter(individual_ID == "M-03a") # but does he have April?? if not, add as a new capture for May (not added rn)
# if he has an Apri msmt, then add 1 recap to May, if not, leve as-is
```

Throughout the summer, we measured 41 collared lizards, & 38 widespread lizards.

Of the 38 widespread lizards, 28 were captured in April, 9 in May, and 1 in July. 5 of the widespread lizards captured in April were recaptured in May. 

Of the 41 collared lizards, 39 were captured (and collared) in April, 22 (of the original 39) were recaptured in May, and 14 (of the original 39) were recaptured in July. F08a and M03a were newly captured in May and radio-collared. F08a was a new capture, but M03a was briefly originally assigned W-029. **check this** *Thankfully, the widespread data no longer includes that data, and it's been shifted to the collared data only.*

How long were lizards held between capture and measurement?

```{r capture to msmt}
# collared
mean(lizards_collared$capture_to_msmt, na.rm = TRUE)/60 # hours
median(lizards_collared$capture_to_msmt, na.rm = TRUE)/60
hist(lizards_collared$capture_to_msmt)
300/60

# widespread
mean(lizards_widespread$capture_to_msmt, na.rm = TRUE) # minutes
max(lizards_widespread$capture_to_msmt, na.rm = TRUE)
```

Most collared lizards were measured within 3 hours of capture, with the majority being measured within 2 hours of capture, and a few being measured the day after capture.

Widespread lizards were always measured within 5 hours of capture, and usually within 2 hours.


How long were treatment lizards in treatment?

```{r tmt time}
mean(lizards_collared$tmt_time_elapsed, na.rm = T)
hist(lizards_collared$tmt_time_elapsed, na.rm = T)
```

Lizards were put in bins with shallow water for 1 hour, on average, to promote drinking behavior.



# Merge dfs

```{r}
merged_data <- lizards_collared %>%
  # remove collared info we only needed for methods summary
  dplyr::select(-treatment_date_time, -tmt_time_elapsed, 
                -radio_collar_mass_g, -radio_collar_percent_mass_g
                ) %>%
  # add widespread lizards
  rbind(lizards_widespread) %>%
  # join CEWL data
  left_join(CEWL, by = c('individual_ID', 
                         'capture_date' = 'date')) %>%
  # join osmolality data
  left_join(osmolality, by = c('individual_ID', 
                               'capture_date' = 'blood_draw_date')) %>%
  # relabel and add factors for informative grouping
  mutate(sex_M_F = factor(sex_M_F, labels = c("Female", "Male")),
         tmt = factor(tmt, levels = c("hydration", "control", "No Tmt"),
                           labels = c("Water Tmt", "Sham Tmt", "No Tmt")),
         month = factor(substr(as.character(capture_date), 7, 7),
                          labels = c("April", "May", "July")),
         week = factor(month, 
                       labels = c("1", "3", "12")),
         week_num = as.numeric(as.character(week))) %>%
  # join female gravidity notes
  left_join(eggs, by = c('individual_ID', 'month'))

summary(merged_data)
```





# Calculate Body Condition

This is also known as scaled mass index, or log-log residuals.

I calculate as described by: Peig, J., & Green, A. J. (2009). New perspectives for estimating body condition from mass/length data: The scaled mass index as an alternative method. Oikos, 118(12), 1883–1891. https://doi.org/10.1111/j.1600-0706.2009.17643.x


Step 0:

get subset of data with only April data:

```{r April only}
April_lizard_dat <- merged_data %>%
  dplyr::filter(month == "April")
```


Step 1: Simple Linear Regression

```{r SMI SLR}
mass_SVL_SLR <- lm(data = April_lizard_dat, mass_g ~ SVL_mm)
summary(mass_SVL_SLR)
```


Step 2: Identify Outliers

```{r SMI equation outliers 1}
plot(mass_SVL_SLR)
```

Hm, residuals are fanned-out, and many have high values, but they are pretty evenly distributed on either side of zero.


So, check for high leverage points:

```{r high leverage}
# compute values for observations 
high_leverage <- data.frame(H = hatvalues(mass_SVL_SLR)) %>% 
  mutate(row = rownames(.))

# compute cutoff value 
h_bar <- (3*sum(high_leverage$H))/nrow(high_leverage)

# add to original dataframe 
# see which observations have extremely high leverage (if any)
high_leverage_dat <- April_lizard_dat %>%
  dplyr::mutate(residuals = residuals(mass_SVL_SLR)) %>%
  dplyr::mutate(row = rownames(.)) %>%
  left_join(., high_leverage, by = "row") %>%
  dplyr::filter(H > h_bar) 
high_leverage_dat
```

The point for individual M-20 seems to be high-leverage, so we will try removing

Check for influential points based on Cook's distance:

```{r}
# get Cook's distance 
cooks <- data.frame(c = cooks.distance(mass_SVL_SLR) # specify model name 
                    ) %>% mutate(row = rownames(.))

# add to original dataframe 
influential <- April_lizard_dat %>%
  dplyr::mutate(residuals = residuals(mass_SVL_SLR)) %>%
  dplyr::mutate(row = rownames(.)) %>% 
  left_join(., cooks, by = "row")

# see moderately influential points 
cook_mod_inf <- influential %>% 
  dplyr::filter(c>0.5) 
cook_mod_inf
```

There are none, so there's nothing to potentially remove.


```{r SMI equation outliers 2}
boxplot(residuals(mass_SVL_SLR))
hist(residuals(mass_SVL_SLR))
max(residuals(mass_SVL_SLR))
influential %>% dplyr::filter(residuals < -10)
influential %>% dplyr::filter(residuals > 8)
```

From the boxplot, there are four individuals with much larger residuals than the rest of the distribution. The histogram looks fine, and pretty normally distributed.


Test whether removing the high-leverage pt and the high residuals improve things:

```{r remove outliers}
cleaned_SMI_dat <- April_lizard_dat %>%
  dplyr::filter(individual_ID %nin% c("W-012", # super low residual
                                      "M-07", "F-03", "M-15", "W-015", # high resids
                                      "M-20" # influential pt
                                      ))
```


re-check residuals:

```{r check 2}
mass_SVL_SLR2 <- lm(data = cleaned_SMI_dat, mass_g ~ SVL_mm)
summary(mass_SVL_SLR2)
plot(mass_SVL_SLR2)
mean(residuals(mass_SVL_SLR2))
median(residuals(mass_SVL_SLR2))
```

I don't think removing those points improved things much with residuals, but R-sq is better, so I will keep the subsetted data.


Step 3: log-log Regression

```{r SMI log log regression}
log_mass_SVL_SLR <- lm(data = cleaned_SMI_dat, 
                       log(mass_g) ~ log(SVL_mm))
summary(log_mass_SVL_SLR)
```


Step 4: Extract Values

compute standardized major axis using the log-log regression equation:

```{r SMI equation}
r <- sqrt(0.8526) # Pearson's correlection coefficient (sqrt of R-squared)
b_OLS <- 2.0891 # regression slope
b_SMA <- b_OLS/r
```

mean length in capture data:

```{r mean SVL}
L0 <- mean(cleaned_SMI_dat$SVL_mm)
```


Step 5: Calculate Scaled Mass Index

```{r join all data}
lizards_all_SMI <- merged_data %>%
  # compute SMI
  mutate(SMI = mass_g * ((L0/SVL_mm) ^ b_SMA))

summary(lizards_all_SMI)
```


# Export Finished Dataframe

```{r}
write_rds(lizards_all_SMI, "./data/G_sila_clean_full_dat.RDS")
```
























